---
title: "Smed CisReg project 02: scRNA Seurat Analysis"
author: "Alberto Perez-Posada @apposada"
date: "6/13/2023"
output:
  html_document: default
  md_document:
    variant: markdown_github
editor_options: 
  markdown: 
    wrap: sentence
---

```{r setup, include=FALSE}
dir <- '/home/alberto/storage/projects/smed_cisreg/'
fcha <- function(){ gsub("-","",Sys.Date()) }

knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = dir)
options(scipen=999)
Sys.setenv(VROOM_CONNECTION_SIZE=5000072)
```

## About

This markdown shows the Seurat Analysis of the single cell RNA-seq dataset.

## Load libraries

```{r libraries, message=FALSE, warning=FALSE}
library(plyr)
library(dplyr)
library(ggplot2)
library(Seurat)
library(harmony)
library(colorspace)
library(RColorBrewer)
```

## Load functions

```{r load_functions, message = FALSE, warning = FALSE}
source("code/r_code/functions/sourcefolder.R")

sourceFolder(
  "code/r_code/functions/", #change for definitive path once its final
  recursive = TRUE
  )
```

## Load data

We start by loading the concatenated matrices we generated in the first markdown:

```{r load_data}
smed_cisreg_scrna <- readRDS("/omics-storage/lab-solana/alberto/projects/smed_cisreg/outputs/rda/20250117_smed_cisreg_scRNA_concatenated_libraries_min100_revision_nogfpi.RDS")
```

## Seurat Analysis

We will carry out a standard Seurat Analysis to identify the clusters of different cell types.

We start by defining highly variable genes. We will choose 20K since we have a very large gene annotation (with 50K independent features)

```{r selectIntegrationFeatures, message = FALSE, warning = FALSE}
var.genes <- 
  SelectIntegrationFeatures(
    SplitObject(smed_cisreg_scrna, split.by = "library"),
    nfeatures = 20000,
    verbose = TRUE,
    fvf.nfeatures = 20000,
    selection.method = "vst"
    )

VariableFeatures(smed_cisreg_scrna) <- var.genes
```

Scale data (center mean around zero and shift deviation towards +/-1). Done prior to PCA:

```{r scaledata}
smed_cisreg_scrna <- ScaleData(smed_cisreg_scrna, features = VariableFeatures(smed_cisreg_scrna))
```

We start by having a look at the basic features of the dataset by library:

```{r basicfeatures, warning=FALSE, message=FALSE, fig.width = 8, fig.height = 6}
BasicFeatures_presubset <- 
  VlnPlot(smed_cisreg_scrna, features = c("nFeature_RNA", "nCount_RNA"))

print(BasicFeatures_presubset)
```

Now we do a PCA on the data including only the variable genes.

```{r pca, message = FALSE, warning = FALSE}
# Do PCA on data including only the variable genes.
smed_cisreg_scrna <- 
  RunPCA(
    smed_cisreg_scrna, 
    features = var.genes,
    npcs = 120, 
    ndims.print = 1:50, 
    nfeatures.print = 5
    )
```

We can visualise a list of the genes that define the top 5 principal components:

```{r pca_plot, message = FALSE, warning = FALSE}
Pca1_2 <- 
  VizDimLoadings(
    smed_cisreg_scrna,
    dims = 1:2,
    reduction = "pca"
    )

print(Pca1_2)
```

Here the plot of the cells scattered throughout the PCA.

```{r}
# Plotted overlapping
dimplot <- 
  DimPlot(
    smed_cisreg_scrna, 
    reduction = "pca", 
    dims = c(1, 2), 
    group.by = "library",
    raster = FALSE
    )
print(dimplot)
```

## Harmony

We run Harmony [ref] to do batch effect correction of all the libraries.

```{r harmony, warning = FALSE, message = FALSE}
smed_cisreg_scrna <- 
  RunHarmony(
    smed_cisreg_scrna, "orig.ident",
    dims.use = 1:120, 
    theta = 3, 
    lambda = 3, 
    nclust = 40,
    max.iter.harmony = 20,
    plot_convergence = TRUE
  )

harmony_embeddings <- Embeddings(smed_cisreg_scrna, 'harmony')
```

## Neighbours and clustering

We will cluster the cells using all of the PCs calculated above. This will embed cells in a knn-graph structure that can be helpful to identify data communities.

```{r findneighbors}
smed_cisreg_scrna <-
  FindNeighbors(
    smed_cisreg_scrna,
    reduction = "harmony",
    dims = 1:120,
    k.param = 35
    )
```

And after finding the neighbors, we find the clusters:

```{r findclusters}
smed_cisreg_scrna <- 
  FindClusters(
    smed_cisreg_scrna,
    resolution = 1.5,
    algorithm = 1,
    random.seed = 75
    )

smed_cisreg_scrna_res15 <- smed_cisreg_scrna$seurat_clusters

smed_cisreg_scrna <- 
  FindClusters(
    smed_cisreg_scrna,
    resolution = 2,
    algorithm = 1,
    random.seed = 75
    )

smed_cisreg_scrna_res2 <- smed_cisreg_scrna$seurat_clusters

smed_cisreg_scrna <- 
  FindClusters(
    smed_cisreg_scrna,
    resolution = 2.5,
    algorithm = 1,
    random.seed = 75
    )

smed_cisreg_scrna_res25 <- smed_cisreg_scrna$seurat_clusters

smed_cisreg_scrna$seurat_clusters_res15 <- smed_cisreg_scrna_res15

smed_cisreg_scrna$seurat_clusters_res2 <- smed_cisreg_scrna_res2

smed_cisreg_scrna$seurat_clusters_res25 <- smed_cisreg_scrna_res25

```

## UMAP Visualisation

Create a few UMAP visualizations. UMAP creates 2 dimensional representations
of the inter-relationships between cells. Cells within clusters will co-localise.

```{r runumap}
smed_cisreg_scrna <- 
  RunUMAP(
    smed_cisreg_scrna, 
    dims = 1:120,
    reduction = "harmony",
    n.neighbors = 45,
    min.dist = 0.5,
    spread = 1,
    metric = "euclidean",
    seed.use = 1
    )
```

Prior to any plotting, we will define a set of colors for the clusters:

```{r colorpalette}
scrna_colors <- 
  c(
    divergingx_hcl(7,"Geyser"),
    divergingx_hcl(7,"TealRose"),
    divergingx_hcl(7,"PRGn"),
    divergingx_hcl(11,"Spectral"),
    divergingx_hcl(7,"Roma"),
    rainbow(6),
    sequential_hcl(7,"Hawaii"),
    sequential_hcl(7,"Batlow"),
    divergingx_hcl(7,"Cividis")
  )

num_clusters <- length(unique(smed_cisreg_scrna$seurat_clusters_res15))

set.seed(123)
cols <- sample(unname(scrna_colors),num_clusters)
```

And finally, the visualisation of the UMAP after library integration using Harmony, with cells from the different libraries in different colors.

```{r umap_group_library, message = FALSE, fig.width = 10, fig.height = 10}
Umap_group_library <- 
  DimPlot(
    smed_cisreg_scrna,
    reduction = "umap",
    group.by = "orig.ident",
    cols = brewer.pal(11,"Spectral"),
    pt.size = 1,
    raster = FALSE
    )

print(Umap_group_library)
```

Here is the UMAP with the clusters:

```{r umap_clusters, fig.height = 10, fig.width = 10}
Umap_cluster <- 
  DimPlot(
    smed_cisreg_scrna, 
    cols = cols,
    reduction = "umap", 
    group.by = "seurat_clusters_res15",
    label = TRUE,
    raster = FALSE
    )

Idents(smed_cisreg_scrna) <- smed_cisreg_scrna$seurat_clusters_res15

print(Umap_cluster)
```

And different UMAPs, clusters by colors and split by library of origin. 

```{r umap_split_library, fig.height = 22, fig.width = 12}
Umap_split_library <- 
  DimPlot(
    smed_cisreg_scrna,
    cols = cols,
    reduction = "umap",
    split.by = "orig.ident",
    ncol = 2,
    raster = FALSE
    )

print(Umap_split_library)
```


We will finally save this file as an object we can load later on:

```{r save}
saveRDS(
  smed_cisreg_scrna,
  file = "outputs/rda/seurat_scrna/smed_cisreg_rna_seuratanalysis_revision_nogfpi.rds"
)
```
