---
title: 'Smed CisReg project 05: scATAC-seq and scRNA-seq combined annotations'
author: "Alberto Perez-Posada @apposada"
date: "7/7/2023"
output: html_document
---

```{r setup, include=FALSE}
dir <- '~/storage/projects/smed_cisreg/'
fcha <- function(){ gsub("-","",Sys.Date()) }

knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = dir)
options(scipen=999)
Sys.setenv(VROOM_CONNECTION_SIZE=5000072)
```

## About

In this markdown we will integrate the scATAC-seq data and the scRNA-seq data.

## Load libraries

```{r libraries, message=FALSE, warning=FALSE}
library(plyr)
library(dplyr)
library(ggplot2)
library(viridis)
library(ComplexHeatmap)
library(Seurat)
library(harmony)
library(colorspace)
library(Signac)
library(GenomicRanges)
library(rtracklayer)
library(ensembldb)
library(ggplot2)
library(igraph)
```

## Load functions

```{r load_functions}
source("code/r_code/functions/sourcefolder.R")

sourceFolder(
  "code/r_code/functions/", #change for definitive path once its final
  recursive = TRUE
  )
```

## Load Data
```{r}
smed_cisreg_scatac <- readRDS("outputs/rda/seurat_scatac/smed_cisreg_scATAC.rds")
```

## Integrating with single-cell transcriptomics data: GeneActivity and TransferLabels

(explain here what gene activity does and the step-by-step to integrate this with the scRNA)

we copied this object over from another server, so we have to update the fragments file.

```{r}
frags <- Fragments(smed_cisreg_scatac)  # get list of fragment objects
Fragments(smed_cisreg_scatac) <- NULL  # remove fragment information from assay

# create a vector with all the new paths, in the correct order for your list of fragment objects
# In this case we only have 1
new.paths <- list("/omics-storage/lab-solana/alberto/projects/smed_cisreg/outputs/scatac/NX73/outs/fragments.tsv.gz")
for (i in seq_along(frags)) {
  frags[[i]] <- UpdatePath(frags[[i]], new.path = new.paths[[i]]) # update path
}

Fragments(smed_cisreg_scatac) <- frags # assign updated list back to the object
Fragments(smed_cisreg_scatac)
```

```{r}
gene.activities <- GeneActivity(smed_cisreg_scatac) # we could also tweak here I think
```

We normalise this data:

```{r}
smed_cisreg_scatac[['RNA']] <- CreateAssayObject(counts = gene.activities)

smed_cisreg_scatac <- NormalizeData(
  object = smed_cisreg_scatac,
  assay = 'RNA',
  normalization.method = 'LogNormalize', # check alternative normalisation method?
  scale.factor = median(smed_cisreg_scatac$nCount_RNA)
)
DefaultAssay(smed_cisreg_scatac) <- "RNA"
```

And much like we did before for the single cell RNA-seq data, we can use that dataset to transfer identity to the cells from this scATAC-experiment. For that we will load the seurat object with the consensus identities assigned.

However, given the lower resolution of the scATAC-seq data, we will use this as reference for knowing which are the broader groups of cell types. This way, plus having the identities from the reference scRNA-seq data (see markdown #03), we can both know the broad cell type of all the clusters of the scRNA-seq data plus the rough identity of the cells in the sc-ATACseq.

```{r}
smed_cisreg_scrna <- readRDS(file = "outputs/rda/seurat_scrna/smed_cisreg_scRNA_consensus_identities_revision_nogfpi.RDS")
smed_cisreg_scrna_ctypes <- 
  read.delim2(
    file = "~/projects/smed_cisreg/outputs/smed_cisreg_ctypes_revision_nogfpi.tsv",
    sep = "\t",
    header = TRUE
  )
```

We confirm that we are working with the same layer of both objects ('RNA')

```{r}
DefaultAssay(smed_cisreg_scatac) == DefaultAssay(smed_cisreg_scrna) #this must be true
```

And then we proceed as before:

```{r}
common_features <- 
  VariableFeatures(smed_cisreg_scrna)[
    VariableFeatures(smed_cisreg_scrna) %in% rownames(smed_cisreg_scatac)
    ]

anchors_scatac_as_ref <- FindTransferAnchors(
  reference = smed_cisreg_scatac,
  query = smed_cisreg_scrna,
  reduction = 'cca', #try lsi
  features = common_features,
  k.anchor = 5,
  k.filter = NA,
  k.score = 10,
  max.features = 1000
)
```

TransferData, we specify the column which corresponds to the scATAC-seq clusters.

```{r}
labels_scatac_as_ref <- TransferData(
  anchorset = anchors_scatac_as_ref,
  refdata = smed_cisreg_scatac$seurat_clusters,
  weight.reduction = smed_cisreg_scrna[['pca']],
  dims = 2:30
)
```

```{r}
smed_cisreg_scrna <- AddMetaData(object = smed_cisreg_scrna, metadata = labels_scatac_as_ref)

smed_cisreg_scrna_scatac_clusters_dimplot <- 
  DimPlot(
    smed_cisreg_scrna,
    reduction = "umap",
    group.by = "predicted.id",
    pt.size = 0.85
  ) + NoAxes()

pdf("graphics/scatac_plots/dimplot_scrna_atacclusters_revision_nogfpi.pdf", width = 6, height = 6)
print(smed_cisreg_scrna_scatac_clusters_dimplot)
dev.off()

print(smed_cisreg_scrna_scatac_clusters_dimplot)
```

There seems to be solid agreement. We will proceed as done in the previous markdown to quantify the labels and see if there is solid consensus.

```{r voting_table}
classifications <- 
  data.frame(
    ctype = smed_cisreg_scrna$ctype,
    atac = smed_cisreg_scrna$predicted.id
  )

classifications_table <- table(classifications$ctype,classifications$atac)
```

This bi-dimensional data can be transformed in a matrix M of ixj clusters and labels respectively, that we can use for visualisation.

```{r M_matrix}
classifications_table <- as.matrix(classifications_table)

M <- matrix(
  classifications_table, 
  nrow = length(unique(classifications$ctype)), 
  ncol = length(unique(classifications$atac)), 
  dimnames = dimnames(classifications_table)
)

M[is.na(M)] <- 0

M <- M[,order(as.numeric(colnames(M)))]

M[1:5,1:5]
```

Since there is a starking difference in cluster size that can make difficult to notice the most frequent label, we can relativise to transform in % (over 1) of cells.

```{r relativise}
M <- t(apply(M,1,relativise))
M[1:5,1:5]
```

And we can visualise this as a heatmap...

```{r heatmap_M, fig.height = 7.5, fig.width = 7.5}
pdf("graphics/scatac_plots/ctypes_transfer_broad_heatmap_revision_nogfpi.pdf")
Heatmap(
  t(M),
  name = "% cells",
  cluster_rows = FALSE,
  col = viridis::viridis(10),
  clustering_method_columns = "complete",
  column_names_side = "top"
)
dev.off()

Heatmap(
  t(M),
  name = "% cells",
  cluster_rows = FALSE,
  col = viridis::viridis(10),
  clustering_method_columns = "complete",
  column_names_side = "top"
)
```

...or we can do something more legible. First we prune of sparse votes (we want hi-confidence candidate labels).

First, we keep transferred labels that have more than 75% popularity in any cluster of our dataset.

```{r prune_M_and_Heatmap, fig.height = 7.5, fig.width = 7.5}
M2 <- M[,apply(M,2,function(x){any(x>0.75)})] # 

pdf("graphics/scatac_plots/ctypes_transfer_broad_heatmap_filt_revision_nogfpi.pdf")
Heatmap(
  t(M2),
  name = "% cells",
  cluster_rows = FALSE,
  col = viridis::viridis(10),
  clustering_method_columns = "complete",
  column_names_side = "top"
)
dev.off()

Heatmap(
  t(M2),
  name = "% cells",
  cluster_rows = FALSE,
  col = viridis(10),
  clustering_method_columns = "complete",
  column_names_side = "top"
)
```

Then, we can transform any value below 75% in 0 to help even more telling them apart.

```{r prune_M2}
M3 <- M2
M3[M3 < 0.75] <- 0
```

And we can turn this into a graph where each of our scRNA cellt ypes is connected to a sc-ATAC cluster. Based on the highest "voted" label, nodes corresponding to labels will connect to the clusters with the highest amount of cells in their cluster.

```{r graph_voting}
g <- graph_from_incidence_matrix(
  incidence = M3,
  directed = TRUE,
  mode = "in",
  weighted = TRUE,
  add.names = NULL
)

# we set up colors
V(g)$col[V(g)$name %in% rownames(M3)] <- 
  translate_ids(
    x=V(g)$name[V(g)$name %in% rownames(M3)],
    dict = smed_cisreg_scrna_ctypes[,c(2,3)]
  )

smed_cisreg_scatac_clu_cols <- setNames(gg_color_hue(12), seq(0,11))

V(g)$col[V(g)$name %in% names(smed_cisreg_scatac_clu_cols)] <- "white"
```

And here the plot:

```{r graph_voting_plot, fig.width = 7.5, fig.height = 7.5}
pdf(
  file = "graphics/scatac_plots/smed_cisreg_scatac_ctypes_transfer_marbles_revision_nogfpi.pdf",
  width = 10,
  height = 10
)
set.seed(123)
plot(
  g,
  vertex.color = V(g)$col,
  vertex.size = 5,
  edge.size = 1,
  edge.arrow.size = 0.2,
  vertex.label.size = 0.1,
  layout = layout_as_bipartite(g)[,c(2,1)]
)
dev.off()

set.seed(123)
plot(
  g,
  vertex.color = V(g)$col,
  vertex.size = 5,
  edge.size = 1,
  edge.arrow.size = 0.5,
  vertex.label.size = 0.1
  )
```

With this information, we can assign broad identities of the clusters in a similar fashion to what we did for the scRNA-seq; and likewise, we can assign broad cell types to the scRNA-seq dataset.

We can also see that the removal of the GFPi libraries did not interfere in the identification of ATAC-Seq clusters, as these associations are identical between the two analyses.
