---
title: 'Smed CisReg project 16: pseudobulk and clustering of OCRs (all)'
author: "Alberto Perez-Posada @apposada"
date: "20/3/2025"
output: html_document
---

```{r setup, include=FALSE}
dir <- '~/storage/projects/smed_cisreg/'
fcha <- function(){ gsub("-","",Sys.Date()) }

knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = dir)
options(scipen=999)
Sys.setenv(VROOM_CONNECTION_SIZE=5000072)
```

## About

In this markdown we will integrate the scATAC-seq data and the scRNA-seq data.

## Load libraries

```{r libraries, message=FALSE, warning=FALSE}
# Libraries
library(circlize)
library(cola)
library(colorspace)
library(ComplexHeatmap)
library(dplyr)
library(EnrichedHeatmap)
library(ensembldb)
library(GenomicRanges)
library(ggplot2)
library(harmony)
library(igraph)
library(plyr)
library(rtracklayer)
library(Seurat)
library(Signac)
library(viridis)
library(DESeq2)
```

## Load functions

```{r load_functions}
source("code/r_code/functions/sourcefolder.R")

sourceFolder(
  "code/r_code/functions/", #change for definitive path once its final
  recursive = TRUE
  )
```

## Load Data
```{r}
smed_cisreg_scatac <- readRDS("outputs/rda/seurat_scatac/smed_cisreg_scATAC_consensus_identities.RDS")

load(file = "outputs/rda/scatac_celltype_utils.rda")
```

## Pseudobulk of the scATAC-Seq

We create here the table of pseudobulk of OCR x cluster.

```{r}
# psbulk
scatac_M <- smed_cisreg_scatac@assays$peaks@counts

scatac_psbulk <- 
  pseudobulk(
    x = scatac_M,
    ident = scatac_idents
  )
```

Here we create the table counting how many cells have every OCR open.

```{r}
cluster_size = 
  c(table(scatac_idents))

scatac_ncells <- 
  pseudobulk_ncells(
    x = scatac_M,
    identities = scatac_idents,
    min_counts = 1
  )
```

## Normalisation using cell weight

We perform the cell weight normalisation that we did for the scRNA-Seq too, please refer to the markdown no.09 for more information.

We will keep track of OCR accessibility underlie by at least 3 cells with 5 counts overall, per cluster, and per OCR.

We apply the size factors normalisation from DESeq as well.

```{r}
min_counts <- 5
min_cells <- 3

scatac_cellweights <- 
  get_cellweight_matrix(
    x = as.matrix(scatac_psbulk),
    y = scatac_ncells,
    C = cluster_size,
    min_counts = min_counts,
    min_cells = min_cells
  )

m_ <- as.matrix(scatac_psbulk)
m_ <- m_[rowSums(m_) >= min_counts,]
m_ <- m_[rownames(m_) %in% rownames(scatac_cellweights),]

m_dds <- DESeqDataSetFromMatrix(
  countData = m_,
  colData = data.frame(condition = colnames(m_)),
  design = ~ condition)
m_dds <- estimateSizeFactors(m_dds)
scatac_counts_norm <- counts(m_dds, normalized=TRUE)

scatac_cw <- log1p(scatac_counts_norm) * scatac_cellweights
```

A quick look at the coefficient of variation of these OCRs. Due to data sparsity, we will keep them all for WGCNA later on.

```{r}
scatac_cw_cv <- apply(scatac_cw,1,function(x){sd(x)/mean(x)})

plot(density(scatac_cw_cv))
abline(v=quantile(scatac_cw_cv), col = rainbow(5))
```

```{r}
pdf("graphics/scatac_cw_boxplot_allpeaks.pdf", he = 5, wi = 5)
boxplot(
  scatac_cw,
  las = 2,
  cex.axis=0.5,
  cex = 0.5,
  col = broad_cols,
  border = darken(broad_cols,.4),
  pch = 16,
  outcol=rgb(0.1,0.1,0.1,0.1),
  main = "cw"
)
dev.off()
```

## Clustering of cell types based on OCR profile similarity

Here is a hierarchical clustering first.

```{r}
scatac_ocrs_all_hcl <- hclust(dist(scatac_cw, method = "euclidean"), method = "ward.D2")
```


```{r, fig.width = 5, fig.height = 8}
### HIERARCHICAL
scatac_hm <-
  Heatmap(
    scatac_cw,
    name = "accessibility",
    cluster_rows = scatac_ocrs_all_hcl,
    clustering_method_columns = "ward.D2",
    show_row_names = FALSE,
    show_column_names = FALSE,
    col = colorRamp2(breaks = c(0,quantile(scatac_cw[scatac_cw>0],.9)), colors = c("#fcfdff", "#174fbc")),
    top_annotation =
      HeatmapAnnotation(
        cluster = colnames(scatac_cw),
        col = list(cluster = broad_cols)
      )
  )
draw(scatac_hm)
```

```{r}
pdf("graphics/scatac_hm_allpeaks_revision.pdf",height = 8, width = 5)
draw(scatac_hm)
dev.off()
```

Co-Occurrence (as Levy et al., 2021)

```{r}
### CO-OCCURENCE
# set fixed seed
set.seed(4343)
h <- c(0.75,0.9)
clustering_algorithm <- "hclust"
clustering_method <- "average"
cor_method <- "pearson"
p <- .25
vargenes = rownames(scatac_cw)

# Levy et al 2021 'treeFromEnsembleClustering' from source
cooc <- treeFromEnsembleClustering(
  x=scatac_cw, p=p, h=h,  n = 1000, vargenes = vargenes, bootstrap=FALSE,
  clustering_algorithm=clustering_algorithm, clustering_method=clustering_method, 
  cor_method=cor_method
)
```

Here we plot this co-occurrence as a heatmap with a tree

```{r, fig.width = 6, fig.height = 5}
row_ha <-
  rowAnnotation(
    cluster = factor(rownames(cooc$cooccurrence), levels = broads),
    col = list( cluster = broad_cols ),
    show_legend = F, show_annotation_name = F
  )

clu_ha <-
  HeatmapAnnotation(
    name = "cell types",
    cluster = factor(colnames(cooc$cooccurrence), levels = broads),
    col = list(cluster = broad_cols),
    show_legend = F, show_annotation_name = F
    )

cooc_col <-
  colorRamp2(
    seq(min(cooc$cooccurrence),max(cooc$cooccurrence),length=9),
    colors = c(
      "white",'#ffffe5','#fff7bc','#fee391','#fec44f',
      '#fe9929','#ec7014','#cc4c02','#990000'
      )
    )

smed_cisreg_cooc_hm <- Heatmap(
  name="co-occurence",
  cooc$cooccurrence,
  col = cooc_col,
  cluster_rows = as.hclust(cooc$tree),
  cluster_columns = as.hclust(cooc$tree),
  left_annotation = row_ha,
  top_annotation = clu_ha,
  row_names_gp = gpar(fontsize = 8),
  column_names_gp = gpar(fontsize = 8)
)

pdf("graphics/scatac_cooc_allpeaks.pdf", wi = 6, he = 5)
draw(smed_cisreg_cooc_hm)
dev.off()

draw(smed_cisreg_cooc_hm)
```

## Save

```{r}
save(
  scatac_psbulk,
  scatac_ncells,
  scatac_cellweights,
  m_dds,
  scatac_counts_norm,
  scatac_cw,
  cooc,
  # cola_cl,
  file = "outputs/rda/scatac_counts/scatac_psbulk_counts_clustering_allpeaks_revision.rda"
)
```