---
title: 'Smed CisReg project 20: ANANSE influence graph analysis'
author: "Alberto Perez-Posada @apposada"
date: "08/05/2024"
output: html_document
---

```{r setup, include=FALSE}
dir <- '~/projects/smed_cisreg/'
fcha <- function(){ gsub("-","",Sys.Date()) }

knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = dir)
options(scipen=999)
Sys.setenv(VROOM_CONNECTION_SIZE=5000072)
```

## About
 include definition of ananse and why we need this etc.

## Load libraries

```{r libraries, message=FALSE, warning=FALSE}
library(circlize)
library(colorspace)
library(ComplexHeatmap)
library(data.table)
library(dplyr)
library(ggplot2)
library(gridExtra)
library(harmony)
library(igraph)
library(plyr)
library(RColorBrewer)
library(reshape2)
library(Seurat)
library(tidyverse)
library(topGO)
library(viridis)
library(WGCNA)
library(xlsx)
```

## Load Functions

```{r load_functions}
source("code/r_code/functions/sourcefolder.R")

sourceFolder(
  "code/r_code/functions/", #change for definitive path once its final
  recursive = TRUE
  )
```

## Load Data

```{r}
load("outputs/rda/tf_analysis.rda")
load("outputs/rda/ananse_dge_vs_neoblast.rda")
load("outputs/rda/ananse_graph_utils.rda")

smed_tfs_m2f <- read.delim2("outputs/ananse/m2f/smed_motif2factors_more_refs_JASPAR_2020/merge_m2f_inference_neiroetal/core_common_tfs/Smed.JASPAR2020.motif2factors.txt")
smed_all_tf_ananse <- unique(smed_tfs_m2f$Factor)

tfs_all <- read.delim2("~/../Standard_References/Smed_Rink/20231127_Rosetta_EXTENDED.tsv")[,c(1,4,5,6,14,30:48)]
tfs_all <- tfs_all[tfs_all$Class != "-",]

ddv6_name <- read.delim2(
  "data/results_ddv6_sequencename.tsv", header = FALSE,
  sep = "\t"
)[,-1]
colnames(ddv6_name) <- c("id_ddv6","sequence_name")

ddv6_name$name <- sub("[^|]+ | ","",ddv6_name$sequence_name)
ddv6_name$name <- sub("\\| ","",ddv6_name$name)
ddv6_name$name <- sub("^ ","",ddv6_name$name)
ddv6_name$name <- gsub("^clone ", "", ddv6_name$name)
ddv6_name$name <- gsub("^Smed[A-Za-z0-9_]+ ", "", ddv6_name$name)
ddv6_name$name <- sub(" \\|  clone [A-Za-z0-9_.]+ unknown mRNA sequence", "", ddv6_name$name)
ddv6_name$name <- gsub(" mRNA.*", "", ddv6_name$name)
ddv6_name$name[grep(" unknown",ddv6_name$name)] <- ""
ddv6_name$name <- sub("strain [A-Za-z0-9\\-]+ ","",ddv6_name$name)
ddv6_name$name <- sub("SMED_[0-9]+_V2","",ddv6_name$name)
ddv6_name$name <- sub("Smed-","",ddv6_name$name)
ddv6_name$name <- gsub("TPA_inf: ", "", ddv6_name$name)
ddv6_name$name <- gsub("UNVERIFIED: ", "", ddv6_name$name)
ddv6_name$name <- gsub("^unknown$","",ddv6_name$name)
ddv6_name$name <- sub("^ ","",ddv6_name$name)

rosetta_ext <- read.delim2("~/../Standard_References/Smed_Rink/20231127_Rosetta_EXTENDED.tsv", header = TRUE)
rosetta_ext$name <-
  paste(
    rosetta_ext$Preferred_name,
    rosetta_ext$Symbol,
    sep = "___"
  )

rosetta_ext$name[grep("-___-", rosetta_ext$name)] <- "-"
rosetta_ext$name[grep("___[A-Za-z0-9]+", rosetta_ext$name, perl = TRUE)] <- gsub(".*___","",rosetta_ext$name[grep("___[A-Za-z0-9]+", rosetta_ext$name, perl = TRUE)])
rosetta_ext$name[grep("___-", rosetta_ext$name, perl = TRUE)] <- gsub("___-", "-like", rosetta_ext$name[grep("___-", rosetta_ext$name, perl = TRUE)])

test <-
  apply(
    rosetta_ext,
    1,
    function(x){
      v6_1 = x[[6]]
      v6_2 = x[[29]]
      v6_2 = unlist(strsplit(v6_2,","))
      
      if(v6_1 %in% ddv6_name$id_ddv6){
        y = ddv6_name$name[ddv6_name$id_ddv6 == v6_1]
      } else {
        y = paste(ddv6_name$name[ddv6_name$id_ddv6 %in% v6_2], collapse = "//")
      }
      return(y)
    }
  )

rosetta_ext$name[which(sapply(test,nchar) != 0)] <- test[which(sapply(test,nchar) != 0)]

rosetta_ext$name_lit <- test
rosetta_ext$name_lit[rosetta_ext$name_lit == ""] <- rosetta_ext$Symbol[rosetta_ext$name_lit == ""]

gene_names <- unique(rosetta_ext[,c(1,49)])

gene_names_lit <- unique(rosetta_ext[,c(1,50)])
gene_names_lit <- gene_names_lit[!(gene_names_lit$name_lit %in% c("","-","//")),]


hnf4 = "h1SMcG0019688"
```


## ANANSE INFLUENCE STUFF

```{r}
broadcols_inf <- broadcols[-c(1,3)] # add or remove epidermis which is no. 4 in this order of colour
names_order_inf <- names(broadcols_inf)
infls <- setNames(paste0(names_order_inf,"_neoblast"), names_order_inf)

infl_dir <- "outputs/ananse/outs/influence/250k/"

# tables of influence
infl_tables <- 
  paste0(
    infl_dir,
    list.files(path = infl_dir, pattern = ".txt")
  )
names(infl_tables) <- gsub(".*infl_","",gsub("_neoblast.*", "", infl_tables))
infl_tables <- infl_tables[match(names_order_inf, names(infl_tables))]

# networks of influence
infl_nws <-
  paste0(
    infl_dir,
    list.files(path = infl_dir, pattern = "_diffnetwork.tsv")
  )
names(infl_nws) <- gsub(".*infl_","",gsub("_neoblast.*", "", infl_nws))
infl_nws <- infl_nws[match(names_order_inf, names(infl_nws))]


lg_inf <- lapply(1:length(infls), function(x) list())
names(lg_inf) <- infls

names(list_DGE_vs_neo)[names(list_DGE_vs_neo) == "BG_neoblast"] <- "basalgoblet_neoblast"
list_DGE_vs_neo <- list_DGE_vs_neo[names(list_DGE_vs_neo) %in% infls]
list_DGE_vs_neo <- list_DGE_vs_neo[match(infls,names(list_DGE_vs_neo))]
```

For the sake of the visualisation, we will be retrieving the top 3 outgoing interactions per TF:

```{r, message = FALSE}
for (i in 1:length(lg_inf)){
  message(i)
  inf <- read.table(infl_tables[i], header = TRUE)
  nw <- read.table(infl_nws[i], header = TRUE)
  g <- graph.data.frame(d = nw, directed = TRUE)
  
  V(g)$centr_score <- nan_to_zero(closeness(g, mode = "out"))
  V(g)$outdegree <- nan_to_zero(igraph::degree(g, mode = "out"))
  
  g_orig <- g
  
  g <- induced_subgraph(g, vids = V(g)$name[V(g)$name %in% inf$factor[1:20]])
  g <- igraph::delete.vertices(g, which(igraph::degree(g) == 0))
  
  fate <- gsub("_neoblast","",names(lg_inf)[i])
  pal <- colorRampPalette(c("#ececec", broadcols_inf[fate]))(5)
  give_col <- circlize::colorRamp2(seq(0,1, len = length(pal)),pal)
  
  V(g)$centr_score <- relativise(V(g)$centr_score)
  V(g)$centr_colour <- give_col(V(g)$centr_score)
  
  V(g)$outdegree <- relativise(V(g)$outdegree)
  V(g)$outdegree_colour <- give_col(V(g)$outdegree)
  
  V(g)$infl_score <- inf$influence_score[match(V(g)$name, inf$factor)]
  V(g)$infl_colour <- give_col(V(g)$infl_score)
  
  # If we want to highlight the D.E. TFs, we can do this:
  V(g)$size <- ifelse(V(g)$name %in% list_DGE_vs_neo[[i]]$diffgenes, 2, .7)
  
  message(i)
  top_edges <- integer()
  for(tf in V(g)$name[igraph::degree(g,V(g), mode = "out")>0]){
    message(tf)
    e <- top_edges_per_tf(g=g, tf = tf, top = 2)
    top_edges <- c(top_edges,e)
  }
  
  g_ <- subgraph.edges(graph = g, eids = E(g)[top_edges], delete.vertices = TRUE)
  
  lg_inf[[i]]$influence <- inf
  lg_inf[[i]]$diff_g <- g_
  lg_inf[[i]]$diff_g_orig <- g_orig
  
  rm(inf,nw,g,g_,e)
}
```

```{r}
library(xlsx)
write.xlsx(
  DF_tgs,
  file = "./outputs/smed_cisreg_ANANSE_target_lists.xlsx",
  sheetName = "target_lists",
  col.names = TRUE, row.names = FALSE, showNA = TRUE
)
```

## Visualising the influence results

```{r}

l_inf = lapply(infl_tables,read.delim2)
l_inf = lapply(l_inf, function(x){})
l_inf = lapply(l_inf, function(x){x$name = translate_ids(x$factor, tfs_all[,c(1,11)]); return(x)})
l_inf$eep$col = broadcols_inf[[1]]
l_inf$epidermis$col = broadcols_inf[[2]]
l_inf$phagocytes$col = broadcols_inf[[3]]
l_inf$basalgoblet$col = broadcols_inf[[4]]
l_inf$muscle$col = broadcols_inf[[5]]
l_inf$neuron$col = broadcols_inf[[6]]
l_inf$parenchyma$col = broadcols_inf[[7]]
l_inf$protonephridia$col = broadcols_inf[[8]]
l_inf$secretory$col = broadcols_inf[[9]]


pdf("graphics/ananse_influence_scatters.pdf",wi = 8, he = 8)
par(mfrow = c(3,3))
for(i in names(l_inf)){
  plot(
    l_inf[[i]]$factor_fc,
    l_inf[[i]]$influence_score,
    pch = 21,
    bg = l_inf[[i]]$col,
    col = "#4c4c4c",
    xlab = "TF FoldChange",
    ylab = "Influence Score",
    main = i
  )
  text(
    x = as.numeric(l_inf[[i]]$factor_fc[1:20]),
    y = as.numeric(l_inf[[i]]$influence_score[1:20]),
    labels = l_inf[[i]]$name[1:20],
    pos = 3,
    cex = 0.6
  )
}
par(mfrow=c(1,1))
dev.off()
```

## Visualising the influence networks

```{r}
pdf("graphics/ananse_diff_networks_symbol_250k_smallsize.pdf", height = 4, width = 4)
for(i in names(lg_inf)){
  set.seed(5678)
  plot(
    main = i,
    lg_inf[[i]]$diff_g,
    layout = layout.graphopt(lg_inf[[i]]$diff_g),
    vertex.color = V(lg_inf[[i]]$diff_g)$outdegree_colour,
    vertex.frame.color = darken(V(lg_inf[[i]]$diff_g)$outdegree_colour,0.5),
    edge.color = rgb(0.1,0.1,0.1,0.25),
    edge.arrow.size = .5,
    vertex.label.family = "Helvetica",
    vertex.label.color = "black",
    vertex.label = translate_ids(V(lg_inf[[i]]$diff_g)$name, tfs_all[,c("gene","Symbol")]),
    vertex.label.cex = .5,
    vertex.size = 8 * V(lg_inf[[i]]$diff_g)$size
  )
}
dev.off()
```

Examples:

```{r}
i = "phagocytes_neoblast"

set.seed(5678)
plot(
  main = i,
  lg_inf[[i]]$diff_g,
  layout = layout.graphopt(lg_inf[[i]]$diff_g),
  vertex.color = V(lg_inf[[i]]$diff_g)$outdegree_colour,
  vertex.frame.color = darken(V(lg_inf[[i]]$diff_g)$outdegree_colour,0.5),
  edge.color = rgb(0.1,0.1,0.1,0.25),
  edge.arrow.size = .5,
  vertex.label.family = "Helvetica",
  vertex.label.color = "black",
  vertex.label = translate_ids(V(lg_inf[[i]]$diff_g)$name, tfs_all[,c("gene","Symbol")]),
  vertex.label.cex = .5,
  vertex.size = 8 * V(lg_inf[[i]]$diff_g)$size
)

```

We can focus in one example to see that there are TFs that have similarly high scores of influence in different clusters:

```{r}
inf_pha_par <-
  merge(
    lg_inf$phagocytes_neoblast$influence[,c(1,2)],
    lg_inf$parenchyma_neoblast$influence[,c(1,2)],
    by = 1,
    all.x = TRUE,
    all.y = TRUE
  )
colnames(inf_pha_par) <- c("id","phagocytes","parenchyma")
rownames(inf_pha_par) <- inf_pha_par$id; inf_pha_par$id <- NULL
inf_pha_par[is.na(inf_pha_par)] <- 0
inf_pha_par <- inf_pha_par[rev(order(inf_pha_par$phagocytes+inf_pha_par$parenchyma)),]
inf_pha_par <- inf_pha_par[inf_pha_par$phagocytes + inf_pha_par$parenchyma > 0,]

colvec <- setNames(rep("#BDC3C7", nrow(inf_pha_par)),rownames(inf_pha_par))
colvec[hnf4] <- "#da9fa0"
colvec["h1SMcG0016896"] <- "#e58bdd"
colvec["h1SMcG0009317"] <- "#66d18a"
dotsizes = setNames(rep(1,nrow(inf_pha_par)), rownames(inf_pha_par))
dotsizes[names(dotsizes) %in% c(hnf4,"h1SMcG0016896","h1SMcG0009317")] <- 2

plot(inf_pha_par, pch = 21, bg = alpha(colvec,.8), col = alpha(darken(colvec,.4),.8), cex = dotsizes, main = "influence of phagocyte and\nparenchymal TFs")

pdf("graphics/ananse_influence_pha_par.pdf", width = 3.5, height = 4)
plot(inf_pha_par, pch = 21, bg = alpha(colvec,.8), col = alpha(darken(colvec,.4),.8), cex = dotsizes, main = "influence of phagocyte and\nparenchymal TFs")
dev.off()

```

## Co-influential TFs

```{r}
tf_inf_l <- 
  lapply(
    lg_inf,
    function(x){
      setNames(x$influence$influence_score,x$influence$factor)
    }
  )
tf_inf <- t(ldply(tf_inf_l,function(s){t(data.frame(unlist(s)))}))
colnames(tf_inf) <- tf_inf[1,]
tf_inf <- as.data.frame(tf_inf[-1,])
tf_inf = data.frame(lapply(tf_inf, function(x) as.numeric(x)),
                    check.names=F, row.names = rownames(tf_inf))
tf_inf[is.na(tf_inf)] = 0

colnames(tf_inf) <- gsub("_neoblast","",colnames(tf_inf))
tf_inf <- tf_inf[,match(names_order_inf, colnames(tf_inf))]

tf_inf <-
  tf_inf[
    match(
      names(sort(sapply(as.data.frame(t(tf_inf)),function(x){which(x == max(x))[1]}))), #staircase sorting
      rownames(tf_inf)
    ),
  ]

tf_inf <- tf_inf[apply(tf_inf,1,function(x){any(x>.3)}),]
```

## Defining clusters of co-influential TFs

```{r}
tf_inf_cor <- cor(t(tf_inf))

set.seed(1234)
tf_inf_clu <- hclust(as.dist(1-tf_inf_cor),method = "complete")
plot(tf_inf_clu)
abline(h=c(.5,.7,.8,.9,1,1.5,2), col = divergingx_hcl(7, "Spectral"), lwd = 1.2)
```

We tidy and arrange the clustering similar to what we have done in WGCNA previously.

```{r}
cut_thresh <- .7

tf_inf_clu_hc <- cutree(tf_inf_clu,h=cut_thresh)

coinf_clusters_table <- 
  reorder_modules(
    data.frame(
      tf_inf,
      module = tf_inf_clu_hc
    ),
    order_criterion = names(broadcols_inf),
    ordering_function = "median",
    thresh_sd = 1 # change for 1.5 if needed
  )

coinf_clusters_table$col <- 
  broadcols_inf[match(coinf_clusters_table$celltypes, names(broadcols_inf))]

coinf_clusters_table$col[grep(",",coinf_clusters_table$celltypes)] <- 
  sapply(
    coinf_clusters_table$celltypes[grep(",",coinf_clusters_table$celltypes)],
    function(x){
      y = average_cols(broadcols_inf[unlist(strsplit(x,split=","))])
      return(y)
    }
  )

if(length(coinf_clusters_table$col[is.na(coinf_clusters_table$col)]) !=0){
  coinf_clusters_table$col[is.na(coinf_clusters_table$col)] <- 
    sapply(
      seq(.1,.6, length = length(coinf_clusters_table$col[is.na(coinf_clusters_table$col)])),
      function(x){y=rgb(x,x,x);return(y)}
    )
}

clu_cols <- 
  setNames(
    coinf_clusters_table$col,
    factor(coinf_clusters_table$module_wgcna, levels = unique(coinf_clusters_table$module_wgcna))
    )

tf_inf_cluID <- setNames(
  translate_ids(tf_inf_clu_hc, coinf_clusters_table[,c(2,4)]),
  names(tf_inf_clu_hc)
  )

head(coinf_clusters_table)
```

Here the overall behaviour of the TFs in these clusters of co-influence

```{r, fig.width = 8, fig.height = 10}
pdf("graphics/ananse_influence_tf_clusters.pdf", height = 14, width = 8)
par(mfrow=c(5,4))
for(i in coinf_clusters_table$module_wgcna){
  boxplot(
    tf_inf[tf_inf_clu_hc == i,],
    ylim = c(0,1),
    las = 2, xaxt = "n", frame.plot = FALSE,
    main = paste0("TF cluster ", i),
    xlab = "celltype",
    ylab = "influence",
    col = broadcols_inf, border = darken(broadcols_inf, .5)
  )
}
par(mfrow = c(1,1))
dev.off()

par(mfrow=c(5,4))
for(i in coinf_clusters_table$module_wgcna){
  boxplot(
    tf_inf[tf_inf_clu_hc == i,],
    ylim = c(0,1),
    las = 2, xaxt = "n", frame.plot = FALSE,
    main = paste0("TF cluster ", i),
    xlab = "celltype",
    ylab = "influence",
    col = broadcols_inf, border = darken(broadcols_inf, .5)
  )
}
par(mfrow = c(1,1))
```

We will extract the top coinfluential TFs using correlation to their own module as a proxy:


```{r}
tf_inf_avg <- aggregate(tf_inf, by = list(clu = tf_inf_cluID), FUN = mean)
rownames(tf_inf_avg) <- tf_inf_avg$clu
tf_inf_avg$clu <- NULL

coinf_cor <- 
  sapply(
    rownames(tf_inf),
    function(x){
      y = tf_inf_cluID[x]
      a = as.numeric(tf_inf[x,])
      b = as.numeric(tf_inf_avg[y,])
      z = cor(a,b)
      return(z)
    }
    )

coinf_cor_df <-
  data.frame(
    id = names(coinf_cor),
    cor = coinf_cor,
    clu = tf_inf_cluID
  )

xlsx::write.xlsx(
  coinf_cor_df,
  file = "./outputs/smed_cisreg_ANANSE_TF_coinfluence_membership.xlsx",
  sheetName = "tf co-influence cluster membership",
  col.names = TRUE, row.names = FALSE, showNA = TRUE
)

set.seed(1234)
coinf_cor_df %>% 
  group_by(clu) %>% 
  slice_max(order_by = cor, n = 5) %>%
  slice_sample(n=5) -> coinf_cor_df_top

# coinf_kme <- cor(t(tf_inf),t(tf_inf_avg)) # I think we do not use this for anything

tf_inf_top <- tf_inf[coinf_cor_df_top$id,]

head(coinf_cor_df_top)
```

Here a heatmap for visualisation:

```{r, fig.width = 2.8, fig.height = 8}
# canijo
tf_inf_hm <-
  Heatmap(
    name = "TF\ninfluence\nscore",
    tf_inf_top,
    col = colorRampPalette(c("#f1f5ff","#b1b8c4","#2c2b46"))(10), # sequential_hcl(10, "Sunset"),
    # show_row_names = FALSE,
    show_column_names = FALSE,
    cluster_rows = FALSE,
    # clustering_method_rows = "ward.D2",
    clustering_method_columns = "ward.D2",
    row_split = factor(coinf_cor_df_top$clu, levels = coinf_clusters_table$newname),
    row_labels = translate_ids(rownames(tf_inf_top),tfs_all[,c(1,11)]),
    row_names_gp = gpar(cex = .7),
    row_title_rot = 0,
    show_row_dend = FALSE,
    row_title_gp = gpar(cex=.7),
    top_annotation = 
      HeatmapAnnotation(
        fate = factor(colnames(tf_inf), levels = colnames(tf_inf)),
        `no. TFs` = anno_barplot(
          sapply(
            lg_inf[match(names_order_inf,gsub("_neoblast","",names(lg_inf)))],
            function(x){nrow(x$influence)}),
          gp = gpar(fill = broadcols_inf, border = darken(broadcols_inf,.5))
        ),
        col = list(fate = setNames(broadcols_inf,colnames(tf_inf))),
        show_legend = FALSE
        ),
    # right_annotation = tf_sym_mark,
    left_annotation = HeatmapAnnotation(
      hc = factor(coinf_cor_df_top$clu, levels = unique(coinf_clusters_table$newname)),
      col = list(
        hc = setNames(clu_cols,coinf_clusters_table$newname[match(names(clu_cols),coinf_clusters_table$module_wgcna)])
      ),
      which = "row",
      show_legend = FALSE
    )
  )

# write.table(
#   merge(
#     coinf_cor_df_top,
#     rosetta_ext,
#     by = 1,
#     all.x = TRUE
#   )[,c(1:8,15,16,29:32,43:46)],
#   file = "~/20240426_TFs_table_COINFLUENCE_CURATION.tsv",
#   sep = "\t",
#   dec = ".",
#   row.names = FALSE,
#   quote = FALSE
# )

pdf("graphics/tf_inf_hm.pdf", width = 2.8, height = 8)
draw(tf_inf_hm)
dev.off()

draw(tf_inf_hm)
```

```{r}
tf_inf_hm_supp <-
  Heatmap(
    name = "TF\ninfluence\nscore",
    tf_inf,
    col = colorRampPalette(c("#f1f5ff","#b1b8c4","#2c2b46"))(10), # sequential_hcl(10, "Sunset"),
    show_row_names = FALSE,
    cluster_columns = TRUE,
    cluster_rows = tf_inf_clu,
    column_names_side = "top",
    show_column_names = FALSE,
    clustering_method_rows = "ward.D2",
    top_annotation = 
      HeatmapAnnotation(
        fate = factor(colnames(tf_inf), levels = colnames(tf_inf)),
        `no. TFs` = anno_barplot(
          sapply(
            lg_inf[match(names_order_inf,gsub("_neoblast","",names(lg_inf)))],
            function(x){nrow(x$influence)}),
          gp = gpar(fill = broadcols_inf, border = darken(broadcols_inf,.5))
        ),
        col = list(fate = setNames(broadcols_inf,colnames(tf_inf))),
        which = "column"
        ),
    left_annotation =
      HeatmapAnnotation(
        hc = factor(tf_inf_clu_hc, levels = unique(coinf_clusters_table$module_wgcna)),
        col = list(
          hc = clu_cols
        ),
        which = "row",
        show_legend = FALSE
      )
  )

pdf("graphics/tf_inf_hm.pdf",width = 4, height = 14)
draw(tf_inf_hm_supp)
dev.off()
```

## Visualising co-influential TFs with igraph

```{r, fig.height = 8, fig.width = 8}
g3 <- graph_from_adjacency_matrix(
  cor(t(tf_inf)),mode = "upper",diag = FALSE, weighted = TRUE
)

g3 <- subgraph.edges(g3, eids = E(g3)[E(g3)$weight > .7])
V(g3)$hc_clu <- tf_inf_clu_hc[match(V(g3)$name, names(tf_inf_clu_hc))]
V(g3)$color_hc <- translate_ids(V(g3)$hc_clu, dict = coinf_clusters_table[,c(2,7)])
ran <- .25+relativise(E(g3)$weight)
ran[ran>1] <- 1
E(g3)$color <- rgb(0.1,0.1,0.1,ran)

pdf("graphics/g3.pdf",width = 8, height = 10)
set.seed(25)
plot(
  g3,
  layout = layout_components(g3),
  vertex.size = 3,
  vertex.label = NA,
  edge.width = E(g3)$weight,
  vertex.color = V(g3)$color_hc,
  vertex.frame.color = darken(V(g3)$color_hc,.4),
  main = "Groups of co-influential TFs"
)
legend(
  'bottomleft',
  legend=coinf_clusters_table$celltypes,
  col= darken(coinf_clusters_table$col, .4),
  pt.bg=coinf_clusters_table$col,pch=21, bty = "n",cex =1
  )
dev.off()

set.seed(25)
plot(
  g3,
  layout = layout_components(g3),
  vertex.size = 3,
  vertex.label = NA,
  edge.width = E(g3)$weight,
  vertex.color = V(g3)$color_hc,
  vertex.frame.color = darken(V(g3)$color_hc,.4),
  main = "Groups of co-influential TFs"
)
legend(
  'bottomleft',
  legend=coinf_clusters_table$celltypes,
  col= darken(coinf_clusters_table$col, .4),
  pt.bg=coinf_clusters_table$col,pch=21, bty = "n",cex =1
  )
```

## Visualising top target genes of each graph:


```{r}
load("outputs/rda/ananse_graph_analysis.rda")
```

```{r}
tfs_inf_stripchart <- read.delim2(
  file = "outputs/functional_annotation/tfs_inf_stripchart.tsv", header = T
)
list_inf_dfs <- lapply(lg_inf, function(x) x$influence)  # Extract data frames
# Create a vector of original sublist names corresponding to each row in data frames
fates <- unlist(lapply(seq_along(list_inf_dfs), function(i) rep(names(lg_inf)[i], nrow(list_inf_dfs[[i]]))))
inf_df_all <- do.call(rbind, list_inf_dfs)
inf_df_all$fate <- fates
inf_df_all$genesymbol <- 
  translate_ids(x = inf_df_all$factor, dict = tfs_all[,c(1,11)])

inf_df_top <- inf_df_all %>% group_by(fate) %>% slice_max(order_by = influence_score, n = 5)
```

Here we retrieve the list of top targets per TF of choice, in their respective graphs:

```{r}
list_tgs <- list()
for(i in unique(inf_df_top$fate)){
  
  fate <- unlist(strsplit(i, "_"))[1]
  
  tfs <- inf_df_top$factor[inf_df_top$fate == i]
  
  list_tgs[[fate]] <- 
    do.call(
      "rbind",
      lapply(
        tfs,
        function(x){
          y = E(lg[[fate]])[.from(x)]$prob # values
          names(y) = head_of(lg[[fate]], es = E(lg[[fate]])[.from(x)])$name # names
          y = y[y > quantile(y, .95)] # filter
          z = data.frame( # make a DF out of this
            fate = fate,
            tf = x,
            gene = names(y),
            value = y,
            value_rel = relativise(y)
          )
          rownames(z) = NULL
          return(z)
        }
      )
    )
}
```

This list we can simply bind together as we dumped all the info required to keep track of each data point (such as source TF, source graph, etc.) in the data frames from the loop right above

```{r}
DF_tgs <- do.call("rbind",list_tgs) # genius

DF_tgs$tf_sym <- translate_ids(x=DF_tgs$tf, dict = tfs_all[,c(1,11)])
DF_tgs$tf_sym[DF_tgs$tf_sym == "-"] <- DF_tgs$tf[DF_tgs$tf_sym == "-"]

DF_tgs$color <-
  broadcols_inf[
    match(
      DF_tgs$fate,
      names(broadcols_inf)
    )
  ]

DF_tgs$fate <- factor(DF_tgs$fate, levels = names(broadcols_inf))

DF_tgs <- DF_tgs[order(DF_tgs$fate, DF_tgs$tf),]

DF_tgs$fate_tf <- paste(as.character(DF_tgs$fate), DF_tgs$tf_sym)
DF_tgs$fate_tf <- factor(DF_tgs$fate_tf, levels = unique(DF_tgs$fate_tf))

DF_tgs$name <- translate_ids(x=DF_tgs$gene, dict = gene_names_lit,return.missing = FALSE)
DF_tgs$name[is.na(DF_tgs$name)] <- ""
head(DF_tgs)


library(xlsx)
write.xlsx(
  DF_tgs,
  file = "./outputs/smed_cisreg_ANANSE_target_lists.xlsx",
  sheetName = "target_lists",
  col.names = FALSE, row.names = TRUE, showNA = TRUE
)
```

This DF can be parsed and used to create lists of plots for each graph separately.

```{r}
ppp <- list()
for(i in levels(DF_tgs$fate)){
  message(i)
  d <- DF_tgs[DF_tgs$fate == i, ]
  
  p_l <- list()
  for(j in unique(d$tf_sym)){
    d_ = d[d$tf_sym == j,]
    
    l_ <- length(unique(d_$name[!(d_$name %in% c(" ","-",""))]))
    
    if(l_ > 30 ){
      set.seed(1234)
      margin_text <- rev(sort(
        sample(unique(d_$name[!(d_$name %in% c(" ","-",""))]),30)
      ))  
    } else{
      margin_text <- rev(sort(unique(d_$name[!(d_$name %in% c(" ","-",""))])))
    }
    
    margin_text <- margin_text[match(d_$name[order(d_$value,decreasing = TRUE)], margin_text)]
    margin_text <- margin_text[complete.cases(margin_text)]
    
    d_$in_text <- ifelse(d_$name %in% margin_text, TRUE, FALSE)
    
    d_ratio_ <- paste0(round(l_ / length(unique(DF_tgs$name)),2)*100,"%")
    
    set.seed(1234)
    p_l[[j]] <- ggplot(d_ %>% arrange(in_text), aes(x = tf_sym, y = value, fill = color, color = in_text, size = in_text)) +
      geom_jitter(width = 0.3, alpha = 0.7, pch = 21) +
      scale_size_manual(values = c(1.5,2.5))+
      scale_fill_identity() +
      scale_color_manual(values = c("white","black"))+
      labs(x = "TF Symbol", y = paste0("Score (top 5% targets) (",d_ratio_," in lit.)")) +
      theme_minimal() +
      theme(plot.margin = unit(c(0, 0, 0, 0), "in"))+
      guides(color ="none", size = "none")+
      grid.text(
        label = margin_text, x = .4, y = seq(0.20, 0.95, length.out = length(margin_text)), 
        gp = gpar(fontsize = 6,fontface = "italic", hjust = 1), draw = FALSE
      )
    
    
  }
  
  ppp[[i]] <- plot_grid(plotlist = p_l, ncol = 5, nrow = 1)
}

pdf("graphics/ananse_top_stripchart.pdf",height = 3.5,width = 12)
for(i in names(ppp)){
  print(ppp[[i]])
}
dev.off()
```

## Saving everything:

```{r}
save(
  lg_inf,
  tf_inf,
  tf_inf_top,
  coinf_clusters_table,
  coinf_cor_df,
  coinf_cor_df_top,
  DF_tgs,
  file = "outputs/rda/ananse_coinfluence.rda"
)

```

```{r}
rosetta_ext_for_paper =
  rosetta_ext

rosetta_ext_for_paper$Symbol[rosetta_ext_for_paper$gene == "h1SMcG0016896"] = "foxf-1"
rosetta_ext_for_paper$name[rosetta_ext_for_paper$gene == "h1SMcG0016896"] = "foxf-1"
rosetta_ext_for_paper$name_lit[rosetta_ext_for_paper$gene == "h1SMcG0016896"] = "foxf-1"

rosetta_ext_for_paper =
  merge(
    rosetta_ext_for_paper,
    smed_id_module,
    by.x = 1,
    by.y = 1,
    all.x = TRUE
  )

write.table(
  rosetta_ext_for_paper,
  "outputs/20240605_supplementary_file_1_1.tsv",
  sep = "\t",
  dec = ".",
  row.names = TRUE,
  quote = FALSE
)
```
