---
title: 'Smed CisReg project 20: ANANSE influence graph analysis. Visualisation of specific networks'
author: "Alberto Perez-Posada @apposada"
date: "08/05/2024"
output: html_document
---

```{r setup, include=FALSE}
dir <- '~/storage/projects/smed_cisreg/'
fcha <- function(){ gsub("-","",Sys.Date()) }

knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = dir)
options(scipen=999)
Sys.setenv(VROOM_CONNECTION_SIZE=5000072)
```

## About

This markdown collects analyses performed in the networks of TF/target genes generated by ANANSE. ANANSE is a tool that leverages gene expression, chromatin accessibility, and TF/motif data to generate networks of putative, in-silico predicted interactions between transcription factors and target genes.

Taking advantage of the resolution of our single cell data, we computationally dissected the gene expression and chromatin accessibility data in this project, to generate profiles of gene expression and chromatin accessibility for each of the major planarian cell types. This allowed us to generate TF/target gene networks for each of these major cell types. By analysing and comparing these networks, we were able to predict putative TFs driving the differentiation from neoblast into each of the major planarian cell fates.

Here we compile all the graph and comparative analyses performed in these networks. We bring to attention the code for visualisation of these networks, and the integration with public data of gene knock-down of TFs in planarians.

Please see the original manuscript, methods, and Supplementary Notes, for more information. Please see also the accompanying markdown to generate the plots of co-influence, and the network plots seen in the supplementary figures. This markdown (the one you're reading right now) is intended to run AFTER THE OTHER ONE.

## Load libraries

```{r libraries, message=FALSE, warning=FALSE}
library(circlize)
library(colorspace)
library(ComplexHeatmap)
library(data.table)
library(dplyr)
library(ggplot2)
library(gridExtra)
library(harmony)
library(igraph)
library(plyr)
library(RColorBrewer)
library(reshape2)
library(Seurat)
library(tidyverse)
library(topGO)
library(viridis)
library(WGCNA)
library(xlsx)
```

## Load Functions

```{r load_functions}
source("code/r_code/functions/sourcefolder.R")

sourceFolder(
  "code/r_code/functions/", #change for definitive path once its final
  recursive = TRUE
  )
```

## Load Data

We will load the TF data, the DGE data for the different broad types (used by ANANSE influence), and some graph utilities we generated.

```{r}
load("outputs/rda/tf_analysis.rda")
load("outputs/rda/ananse_dge_vs_neoblast.rda")
load("outputs/rda/ananse_graph_utils.rda")
```

We generate a modified version of our Smed annotation table containing the gene names (from planmine):

```{r}
tfs_all <- read.delim2("~/storage/../standard_references/Smed_Rink/20231127_Rosetta_EXTENDED.tsv")[,c(1,4,5,6,14,30:48)]
tfs_all <- tfs_all[tfs_all$Class != "-",]

ddv6_name <- read.delim2(
  "data/results_ddv6_sequencename.tsv", header = FALSE,
  sep = "\t"
)[,-1]
colnames(ddv6_name) <- c("id_ddv6","sequence_name")

ddv6_name$name <- sub("[^|]+ | ","",ddv6_name$sequence_name)
ddv6_name$name <- sub("\\| ","",ddv6_name$name)
ddv6_name$name <- sub("^ ","",ddv6_name$name)
ddv6_name$name <- gsub("^clone ", "", ddv6_name$name)
ddv6_name$name <- gsub("^Smed[A-Za-z0-9_]+ ", "", ddv6_name$name)
ddv6_name$name <- sub(" \\|  clone [A-Za-z0-9_.]+ unknown mRNA sequence", "", ddv6_name$name)
ddv6_name$name <- gsub(" mRNA.*", "", ddv6_name$name)
ddv6_name$name[grep(" unknown",ddv6_name$name)] <- ""
ddv6_name$name <- sub("strain [A-Za-z0-9\\-]+ ","",ddv6_name$name)
ddv6_name$name <- sub("SMED_[0-9]+_V2","",ddv6_name$name)
ddv6_name$name <- sub("Smed-","",ddv6_name$name)
ddv6_name$name <- gsub("TPA_inf: ", "", ddv6_name$name)
ddv6_name$name <- gsub("UNVERIFIED: ", "", ddv6_name$name)
ddv6_name$name <- gsub("^unknown$","",ddv6_name$name)
ddv6_name$name <- sub("^ ","",ddv6_name$name)

rosetta_ext <- read.delim2("~/storage/../standard_references/Smed_Rink/20231127_Rosetta_EXTENDED.tsv", header = TRUE)
rosetta_ext$name <-
  paste(
    rosetta_ext$Preferred_name,
    rosetta_ext$Symbol,
    sep = "___"
  )

rosetta_ext$name[grep("-___-", rosetta_ext$name)] <- "-"
rosetta_ext$name[grep("___[A-Za-z0-9]+", rosetta_ext$name, perl = TRUE)] <- gsub(".*___","",rosetta_ext$name[grep("___[A-Za-z0-9]+", rosetta_ext$name, perl = TRUE)])
rosetta_ext$name[grep("___-", rosetta_ext$name, perl = TRUE)] <- gsub("___-", "-like", rosetta_ext$name[grep("___-", rosetta_ext$name, perl = TRUE)])

dd_with_names <-
  apply(
    rosetta_ext,
    1,
    function(x){
      v6_1 = x[[6]]
      v6_2 = x[[29]]
      v6_2 = unlist(strsplit(v6_2,","))
      
      if(v6_1 %in% ddv6_name$id_ddv6){
        y = ddv6_name$name[ddv6_name$id_ddv6 == v6_1]
      } else {
        y = paste(ddv6_name$name[ddv6_name$id_ddv6 %in% v6_2], collapse = "//")
      }
      return(y)
    }
  )

rosetta_ext$name[which(sapply(dd_with_names,nchar) != 0)] <- dd_with_names[which(sapply(dd_with_names,nchar) != 0)]

rosetta_ext$name_lit <- dd_with_names
rosetta_ext$name_lit[rosetta_ext$name_lit == ""] <- rosetta_ext$Symbol[rosetta_ext$name_lit == ""]

gene_names <- unique(rosetta_ext[,c(1,49)])

gene_names_lit <- unique(rosetta_ext[,c(1,50)])
gene_names_lit <- gene_names_lit[!(gene_names_lit$name_lit %in% c("","-","//")),]


hnf4 = "h1SMcG0019688"
```


## Preparation of ANANSE influence data

We will load here the list of influence tables, and prepare some objects (lists, vectors) that will be of use when porting the networks files over.

```{r}
broadcols_inf <- broadcols[-c(1,3)] # add or remove epidermis which is no. 4 in this order of colour
names_order_inf <- names(broadcols_inf)
infls <- setNames(paste0(names_order_inf,"_neoblast"), names_order_inf)

infl_dir <- "outputs/ananse/outs/influence/250k/"

# tables of influence
infl_tables <- 
  paste0(
    infl_dir,
    list.files(path = infl_dir, pattern = ".txt")
  )
names(infl_tables) <- gsub(".*infl_","",gsub("_neoblast.*", "", infl_tables))
infl_tables <- infl_tables[match(names_order_inf, names(infl_tables))]

# networks of influence
infl_nws <-
  paste0(
    infl_dir,
    list.files(path = infl_dir, pattern = "_diffnetwork.tsv")
  )
names(infl_nws) <- gsub(".*infl_","",gsub("_neoblast.*", "", infl_nws))
infl_nws <- infl_nws[match(names_order_inf, names(infl_nws))]


lg_inf <- lapply(1:length(infls), function(x) list())
names(lg_inf) <- infls

names(list_DGE_vs_neo)[names(list_DGE_vs_neo) == "BG_neoblast"] <- "basalgoblet_neoblast"
list_DGE_vs_neo <- list_DGE_vs_neo[names(list_DGE_vs_neo) %in% infls]
list_DGE_vs_neo <- list_DGE_vs_neo[match(infls,names(list_DGE_vs_neo))]
```

## Loading the ANANSE influence networks. Subsetting.

For the sake of the visualisation, we will be retrieving the top 3 outgoing interactions per TF:

```{r, message = FALSE}
for (i in 1:length(lg_inf)){
  message(i)
  inf <- read.table(infl_tables[i], header = TRUE)
  nw <- read.table(infl_nws[i], header = TRUE)
  g <- graph.data.frame(d = nw, directed = TRUE)
  
  V(g)$centr_score <- nan_to_zero(closeness(g, mode = "out"))
  V(g)$outdegree <- nan_to_zero(igraph::degree(g, mode = "out"))
  
  g_orig <- g
  
  g <- induced_subgraph(g, vids = V(g)$name[V(g)$name %in% inf$factor[1:20]])
  g <- igraph::delete.vertices(g, which(igraph::degree(g) == 0))
  
  fate <- gsub("_neoblast","",names(lg_inf)[i])
  pal <- colorRampPalette(c("#ececec", broadcols_inf[fate]))(5)
  give_col <- circlize::colorRamp2(seq(0,1, len = length(pal)),pal)
  
  V(g)$centr_score <- relativise(V(g)$centr_score)
  V(g)$centr_colour <- give_col(V(g)$centr_score)
  
  V(g)$outdegree <- relativise(V(g)$outdegree)
  V(g)$outdegree_colour <- give_col(V(g)$outdegree)
  
  V(g)$infl_score <- inf$influence_score[match(V(g)$name, inf$factor)]
  V(g)$infl_colour <- give_col(V(g)$infl_score)
  
  # If we want to highlight the D.E. TFs, we can do this:
  V(g)$size <- ifelse(V(g)$name %in% list_DGE_vs_neo[[i]]$diffgenes, 2, .7)
  
  message(i)
  top_edges <- integer()
  for(tf in V(g)$name[igraph::degree(g,V(g), mode = "out")>0]){
    message(tf)
    e <- top_edges_per_tf(g=g, tf = tf, top = 2)
    top_edges <- c(top_edges,e)
  }
  
  g_ <- subgraph.edges(graph = g, eids = E(g)[top_edges], delete.vertices = TRUE)
  
  lg_inf[[i]]$influence <- inf
  lg_inf[[i]]$diff_g <- g_
  lg_inf[[i]]$diff_g_orig <- g_orig
  
  rm(inf,nw,g,g_,e)
}
```

## Visualising the influence networks of specific TFs

In the previous markdown we visualised a subset of the influential networks. In this markdown we will visualise the networks of several transcription factors of interest that have been previously studied in the literature. We have functional knockdown data (bulk RNA-Seq) from these, which means that we can dig the validity of our networks by checking the position of the DEGs in the network (as having higher interactions etc.).

We aim at showing the TF of interest and their top interactions with other TFs. Color indicates influence on the cell type. If the surrounding network of the TF of interest is very dense (more than 50 interactions), we subset to keep only interactions that are even more top, followed by some high interactions between the neighbour genes only.

```{r}
lg_functional <- list()

prep = "h1SMcG0004695"
pax5 = "h1SMcG0008441"
soxp = "h1SMcG0009867"
p53 = "h1SMcG0016868"
coe = "h1SMcG0016003"


tf_fate <-
  list(
    coe = c(coe,"neuron_neoblast"),
    pax5 = c(pax5,"epidermis_neoblast"),
    p53 = c(p53,"eep_neoblast"),
    soxp = c(soxp,"eep_neoblast")
  )

for(i in names(tf_fate)){
  tf <- tf_fate[[i]][1]
  fate <- tf_fate[[i]][2]
  inf <- lg_inf[[fate]]$influence
  
  g_f <- lg_inf[[fate]]$diff_g_orig
  g_f <- induced_subgraph(g_f, vids = which(V(g_f)$name %in% V(subgraph_from_edges(g_f, E(g_f)[.from(tf)]))$name ))
  g_f <- induced_subgraph(g_f, which(V(g_f)$name %in% smed_tfs$id))
  g_f <- subgraph_from_edges(g_f, which(E(g_f)$weight > 0.5 ))
  
  if( ecount(g_f) > 50 ){
    
    # top 20 from the one of interest
    g2 <- subgraph_from_edges(g_f, eids = E(g_f)[.from(tf)] )
    nn <- min(ecount(g2),50)
    g2 <- subgraph_from_edges(g2, eids = which(E(g2)$weight >= rev(sort(E(g2)$weight))[nn]) )
    
    # some high ones from others?
    g3 <- induced_subgraph(g_f, vids = which(V(g_f)$name %in% V(g2)$name),impl = "create_from_scratch" )
    g3 <- subgraph_with_top_edges_per_tf(g3,top = 3, delete_isolated = TRUE,mode = "both")
    g3 <- delete.vertices(g3, v = V(g3)$name[V(g3)$name %in% tf])
    
    attrs <- rbind(as_data_frame(g2, "vertices"), as_data_frame(g3, "vertices")) %>% unique()
    el <- rbind(as_data_frame(g2), as_data_frame(g3)) %>% unique()
    new_g <- graph_from_data_frame(el, directed = TRUE, vertices = attrs)
    
    new_g <- induced_subgraph(new_g, vids = which(igraph::degree(new_g,V(new_g), mode = "all") > 1),impl = "create_from_scratch" )
    
    g_f <- new_g
    
    rm(g2,g3,new_g,nn)
    
  } 
  
  pal <- colorRampPalette(c("#ececec", broadcols_inf[gsub("_neoblast","",fate)]))(5)
  give_col <- circlize::colorRamp2(seq(0,1, len = length(pal)),pal)
  
  V(g_f)$infl_score <- inf$influence_score[match(V(g_f)$name, inf$factor)]
  V(g_f)$infl_colour <- give_col(V(g_f)$infl_score)
  V(g_f)$infl_colour[is.na(V(g_f)$infl_colour)] <- "#ececec"
  
  lg_functional[[i]] <- g_f
  message(i)
}
```

Examples:

```{r, fig.width = 4, fig.height = 4}
for(i in names(lg_functional)){
  set.seed(5678)
  plot(
    main = i,
    lg_functional[[i]],
    layout = layout_nicely(lg_functional[[i]]),
    vertex.color = V(lg_functional[[i]])$infl_colour,
    vertex.frame.color = darken(V(lg_functional[[i]])$infl_colour,0.5),
    edge.color = rgb(0.1,0.1,0.1,0.25),
    edge.arrow.size = .5,
    vertex.label.family = "Helvetica",
    vertex.label.color = "black",
    vertex.label = translate_ids(V(lg_functional[[i]])$name, tfs_all[,c("gene","Symbol")]),
    vertex.label.cex = .5,
    vertex.size = 12 #* V(lg_functional[[i]])$size
  )
}
```

```{r}
pdf("graphics/ananse_networks_tfs_functional_summarised.pdf", height = 4, width = 4)
for(i in names(lg_functional)){
  set.seed(5678)
  plot(
    main = i,
    lg_functional[[i]],
    layout = layout_nicely(lg_functional[[i]]),
    vertex.color = V(lg_functional[[i]])$infl_colour,
    vertex.frame.color = darken(V(lg_functional[[i]])$infl_colour,0.5),
    edge.color = rgb(0.1,0.1,0.1,0.25),
    edge.arrow.size = .5,
    vertex.label.family = "Helvetica",
    vertex.label.color = "black",
    vertex.label = translate_ids(V(lg_functional[[i]])$name, tfs_all[,c("gene","Symbol")]),
    vertex.label.cex = .5,
    vertex.size = 12 #* V(lg_functional[[i]])$size
  )
}
dev.off()
```

Please refer to the re-analysis markdowns to see the associated boxplots of interactions between these TFs and the down-regulated genes detected in those previous works in the literature.

## Showing networks of co-influential TFs

We are going to visualise now the network of HNF4 and alx3, two transcription factors that were detected as highly co-influential of two major groups of cell types.

```{r}
hnf4 <- "h1SMcG0019688"
alx3 <- "h1SMcG0021040"

lg_coinf <- list()

coinf_fate <-
  list(
    h_pha = c(hnf4,"phagocytes_neoblast"),
    h_par = c(hnf4,"parenchyma_neoblast"),
    h_bg = c(hnf4,"basalgoblet_neoblast"),
    a_neu = c(alx3,"neuron_neoblast"),
    a_mus = c(alx3,"muscle_neoblast"),
    a_sec = c(alx3,"secretory_neoblast")
  )

```

Here we create these graphs following exactly the same criterion as before.

```{r}
for(i in names(coinf_fate)){
  tf <- coinf_fate[[i]][1]
  fate <- coinf_fate[[i]][2]
  inf <- lg_inf[[fate]]$influence
  
  g_f <- lg_inf[[fate]]$diff_g_orig
  g_f <- induced_subgraph(g_f, vids = which(V(g_f)$name %in% V(subgraph_from_edges(g_f, E(g_f)[.from(tf)]))$name ))
  g_f <- induced_subgraph(g_f, which(V(g_f)$name %in% smed_tfs$id))
  g_f <- subgraph_from_edges(g_f, which(E(g_f)$weight > 0.5 ))
  
  if( ecount(g_f) > 30 ){
    
    # v_add <- E()
    # top 20 from the one of interest
    g2 <- subgraph_from_edges(g_f, eids = E(g_f)[.from(tf)] )
    nn <- min(ecount(g2),30)
    g2 <- subgraph_from_edges(g2, eids = which(E(g2)$weight >= rev(sort(E(g2)$weight))[nn]) )
    
    # some high ones from others?
    g3 <- induced_subgraph(g_f, vids = which(V(g_f)$name %in% V(g2)$name),impl = "create_from_scratch" )
    g3 <- subgraph_with_top_edges_per_tf(g3,top = 3, delete_isolated = TRUE,mode = "both")
    g3 <- delete.vertices(g3, v = V(g3)$name[V(g3)$name %in% tf])
    
    attrs <- rbind(as_data_frame(g2, "vertices"), as_data_frame(g3, "vertices")) %>% unique()
    el <- rbind(as_data_frame(g2), as_data_frame(g3)) %>% unique()
    new_g <- graph_from_data_frame(el, directed = TRUE, vertices = attrs)
    
    g_f <- new_g
    
    rm(g2,g3,new_g,nn)
    
  } 
  
  pal <- colorRampPalette(c("#ececec", broadcols_inf[gsub("_neoblast","",fate)]))(5)
  give_col <- circlize::colorRamp2(seq(0,1, len = length(pal)),pal)
  
  V(g_f)$infl_score <- inf$influence_score[match(V(g_f)$name, inf$factor)]
  V(g_f)$infl_colour <- give_col(V(g_f)$infl_score)
  V(g_f)$infl_colour[is.na(V(g_f)$infl_colour)] <- "#ececec"
  
  lg_coinf[[i]] <- g_f
  message(i)
}
```

Here's the plot:

```{r, fig.width = 6, fig.height = 6}
for(i in names(lg_coinf)){
  set.seed(5678)
  plot(
    main = i,
    lg_coinf[[i]],
    layout = layout.graphopt(lg_coinf[[i]]),
    vertex.color = V(lg_coinf[[i]])$infl_colour,
    vertex.frame.color = darken(V(lg_coinf[[i]])$infl_colour,0.5),
    edge.color = rgb(0.1,0.1,0.1,0.25),
    edge.arrow.size = .5,
    vertex.label.family = "Helvetica",
    vertex.label.color = "black",
    vertex.label = translate_ids(V(lg_coinf[[i]])$name, tfs_all[,c("gene","Symbol")]),
    vertex.label.cex = .5,
    vertex.size = 8 #* V(lg_coinf[[i]])$size
  )
}
```

```{r}
pdf("graphics/ananse_networks_hnf_alx.pdf", wi = 5, he = 5)
for(i in names(lg_coinf)){
  set.seed(5678)
  plot(
    main = i,
    lg_coinf[[i]],
    layout = layout.graphopt(lg_coinf[[i]]),
    vertex.color = V(lg_coinf[[i]])$infl_colour,
    vertex.frame.color = darken(V(lg_coinf[[i]])$infl_colour,0.5),
    edge.color = rgb(0.1,0.1,0.1,0.25),
    edge.arrow.size = .5,
    vertex.label.family = "Helvetica",
    vertex.label.color = "black",
    vertex.label = translate_ids(V(lg_coinf[[i]])$name, tfs_all[,c("gene","Symbol")]),
    vertex.label.cex = .5,
    vertex.size = 8 #* V(lg_coinf[[i]])$size
  )
}
dev.off()
```

## The End!


```{r}
sessionInfo()
```

