---
title: "Smed CisReg project 23: Planarian HNF4 knock-down and Differential Gene Expression; Seurat Analysis"
author: "Alberto Perez-Posada @apposada"
date: "8/11/2023"
output:
  html_document: default
  md_document:
    variant: markdown_github
editor_options: 
  markdown: 
    wrap: sentence
---

```{r setup, include=FALSE}

dir <- '/home/alberto/storage/projects/smed_cisreg/'
fcha <- function(){ gsub("-","",Sys.Date()) }

knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = dir)
options(scipen=999)
Sys.setenv(VROOM_CONNECTION_SIZE=5000072)

```

## About

In this markdown we will assess the differences in cell abundances in the *hnf4* knockdown single cell transcriptomics dataset. For that we will quantify the number of cells per cluster, condition and replicate, and have it assessed by a statistical test such as Chi-Squared.

## Load libraries

```{r libraries, message=FALSE, warning=FALSE}
library(vroom)
library(plyr)
library(dplyr)
library(ggplot2)
library(Seurat)
library(Matrix)
library(harmony)
library(colorspace)
library(RColorBrewer)
```

## Load functions

```{r load_functions}
source("code/r_code/functions/sourcefolder.R")

sourceFolder(
  "code/r_code/functions/", #change for definitive path once its final
  recursive = TRUE
  )
```


## Load data

```{r}
smed_hnf_scrna <- readRDS("outputs/rda/seurat_scrna/hnf4_rnai/smed_hnf_scrna_SeuratObject.RDS")
```

## Umaps for replicates and experiments separately

First we will visualise the position of the cells in the UMAP as a simple guidance.

```{r, fig.width = 8, fig.height = 8}
gfpi_b1 <- smed_hnf_scrna$experiment == "control" & smed_hnf_scrna$replicate == "B1"
gfpi_b2 <- smed_hnf_scrna$experiment == "control" & smed_hnf_scrna$replicate == "B2"
hnfi_b1 <- smed_hnf_scrna$experiment == "HNF4i" & smed_hnf_scrna$replicate == "B1"
hnfi_b2 <- smed_hnf_scrna$experiment == "HNF4i" & smed_hnf_scrna$replicate == "B2"

smed_hnf_scrna$gfpi_b1 <- ifelse(gfpi_b1, "1","0")
smed_hnf_scrna$gfpi_b2 <- ifelse(gfpi_b2, "1","0")
smed_hnf_scrna$hnfi_b1 <- ifelse(hnfi_b1, "1","0")
smed_hnf_scrna$hnfi_b2 <- ifelse(hnfi_b2, "1","0")

gfpi_b1_col <- setNames(ifelse(gfpi_b1, gb_col["orange1"],"#ececec"),smed_hnf_scrna$gfpi_b1)
gfpi_b2_col <- setNames(ifelse(gfpi_b2, gb_col["orange2"],"#ececec"),smed_hnf_scrna$gfpi_b2)
hnfi_b1_col <- setNames(ifelse(hnfi_b1, gb_col["purple1"],"#ececec"),smed_hnf_scrna$hnfi_b1)
hnfi_b2_col <- setNames(ifelse(hnfi_b2, gb_col["blue2"],"#ececec"),smed_hnf_scrna$hnfi_b2)


a1 <- DimPlot(
    smed_hnf_scrna,
    reduction = "umap",
    group.by = "gfpi_b1",
    cols = gfpi_b1_col,
    order = TRUE,
    pt.size = 1
    )+NoAxes()+NoLegend()+labs(title = "Control B1") 

a2 <- DimPlot(
    smed_hnf_scrna,
    reduction = "umap",
    group.by = "gfpi_b2",
    cols = gfpi_b2_col,
    order = TRUE,
    pt.size = 1
    )+NoAxes()+NoLegend()+labs(title = "Control B2") 

b1 <- DimPlot(
    smed_hnf_scrna,
    reduction = "umap",
    group.by = "hnfi_b1",
    cols = hnfi_b1_col,
    order = TRUE,
    pt.size = 1
    )+NoAxes()+NoLegend()+labs(title = "HNF4i B1") 

b2 <- DimPlot(
    smed_hnf_scrna,
    reduction = "umap",
    group.by = "hnfi_b2",
    cols = hnfi_b2_col,
    order = TRUE,
    pt.size = 1
    )+NoAxes()+NoLegend()+labs(title = "HNF4i B2") 

print(
  a1+a2+b1+b2
)

# png("graphics/smed_hnf_umaps_byexp_byrep.png",height = 1200, width = 1200)
# print(
#   a1+a2+b1+b2
# )
# dev.off()
```

In addition, just for good measure, we will see where *hnf4* is being expressed in these cells.

```{r}
smed_hnf_scrna$experiment <- factor(smed_hnf_scrna$experiment,levels = c("control","HNF4i"))
p <-
  FeaturePlot(
    smed_hnf_scrna,
    features = "h1SMcG0019688",
    # cols = c("lightgrey","#a100d2","#4c1f89"),
    split.by = "experiment",
    pt.size = 0.8,
    order = T,
    label = FALSE,
    slot = "counts"
  ) & scale_colour_gradientn(colours = colorRampPalette(c("lightgrey","#a100d2","#4c1f89"))(11)) & 
  theme(legend.position = c(0.1,0.2))

print(p)

png(
  file = "graphics/hnf4_by_experiment.png",
  width= 512,
  height = 256
)
plot_grid(plotlist = lapply(p,function(x){x+NoAxes()}),ncol = 2)
dev.off()
```


## Barplot of percentage of cells per cluster

Next comes a barplot that can help more visually to see where the differences are.

```{r, fig.height = 6, fig.width = 8}
smed_hnf_ctypes <- 
  unique(
  data.frame(
    cluster = smed_hnf_scrna$seurat_clusters,
    ctype = smed_hnf_scrna$ctype,
    broadtype = smed_hnf_scrna$broadtype
    )
  )

smed_ctypes <- read.delim2("outputs/smed_cisreg_ctypes.tsv",header =TRUE)

smed_hnf_ctypes$col <- 
  translate_ids(
    smed_hnf_ctypes$ctype,
    dict = smed_ctypes[,c(1,3)]
  )

smed_hnf_ctypes$col[smed_hnf_ctypes$ctype == "aberrant phagocyte progenitors"] <- "#98b993"

smed_hnf_ctypes$broacol <- 
  translate_ids(
    smed_hnf_ctypes$broadtype,
    dict = unique(smed_ctypes[,c(4,5)])
  )

smed_hnf_scrna_ctypes <- smed_hnf_ctypes
  
smed_hnf_scrna_ctypes_2 <- smed_hnf_scrna_ctypes
smed_hnf_scrna_ctypes_2 <- smed_hnf_scrna_ctypes_2[order(smed_hnf_scrna_ctypes_2$ctype,as.numeric(smed_hnf_scrna_ctypes_2$cluster)),]
smed_hnf_scrna_ctypes_2$cluster <- factor(smed_hnf_scrna_ctypes_2$cluster, levels = unique(smed_hnf_scrna_ctypes_2$cluster))
smed_hnf_scrna_ctypes_2$clu_maxID <- 
  paste(as.character(smed_hnf_scrna_ctypes_2$cluster), as.character(smed_hnf_scrna_ctypes_2$ctype), sep = "_")
smed_hnf_scrna_ctypes_2$clu_maxID <- factor(smed_hnf_scrna_ctypes_2$clu_maxID, levels = unique(smed_hnf_scrna_ctypes_2$clu_maxID))

smed_hnf_scrna$clu_maxID <- translate_ids(smed_hnf_scrna$seurat_clusters, dict = smed_hnf_scrna_ctypes_2[,c(1,6)])

smed_hnf_scrna$exp_rep <- paste(smed_hnf_scrna$experiment,smed_hnf_scrna$replicate, sep = "_")
smed_hnf_scrna$exp_rep <- factor(smed_hnf_scrna$exp_rep, levels = sort(unique(smed_hnf_scrna$exp_rep)))

DF <- 
  data.frame(
    cell = colnames(smed_hnf_scrna),
    experiment = smed_hnf_scrna$experiment,
    replicate = smed_hnf_scrna$replicate,
    exp_rep = smed_hnf_scrna$exp_rep,
    clu_maxID = smed_hnf_scrna$clu_maxID,
    cluster = factor(smed_hnf_scrna$seurat_clusters, levels = levels(smed_hnf_scrna_ctypes_2$cluster))
  )

df <- 
  data.frame(
    table(
      DF$exp_rep, DF$cluster
      )
    )

# Rename the columns of the data frame
colnames(df) <- 
  c("Sample", "Cluster", "Count")

# Normalize the counts in each cluster
df$CountNorm <- df$Count / tapply(df$Count, df$Cluster, sum)[df$Cluster]

exp_cols <- setNames(gb_col[c("orange1","orange2","purple1","blue2")],levels(df$Sample))#setNames(c("#ee875a", "#e87656", "#6BA995", "#43a198"),levels(df$Sample))

# Create a stacked barplot norm number
normncells_bar <- ggplot(df, aes(x = Cluster, y = CountNorm, fill = Sample)) +
  geom_col(position = "stack") +
  scale_fill_manual(values = exp_cols) +
  theme_minimal()+
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  ylab("Proportion of cells")

# Just replicate 1
df_B1 <- 
  data.frame(
    table(
      DF$exp_rep[DF$replicate=="B1"], DF$cluster[DF$replicate=="B1"]
      )
    )

# Rename the columns of the data frame
colnames(df_B1) <- 
  c("Sample", "Cluster", "Count")

# Normalize the counts in each cluster
df_B1$CountNorm <- df_B1$Count / tapply(df_B1$Count, df_B1$Cluster, sum)[df_B1$Cluster]

# Create a stacked barplot norm number
normncells_bar_B1 <- ggplot(df_B1, aes(x = Cluster, y = CountNorm, fill = Sample)) +
  geom_col(position = "stack") +
  scale_fill_manual(values = exp_cols) +
  theme_minimal()+
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  ylab("Proportion of cells")

# Just replicate 2
df_B2 <- 
  data.frame(
    table(
      DF$exp_rep[DF$replicate=="B2"], DF$cluster[DF$replicate=="B2"]
      )
    )

# Rename the columns of the data frame
colnames(df_B2) <- 
  c("Sample", "Cluster", "Count")

# Normalize the counts in each cluster
df_B2$CountNorm <- df_B2$Count / tapply(df_B2$Count, df_B2$Cluster, sum)[df_B2$Cluster]

# Create a stacked barplot norm number
normncells_bar_B2 <- ggplot(df_B2, aes(x = Cluster, y = CountNorm, fill = Sample)) +
  geom_col(position = "stack") +
  scale_fill_manual(values = exp_cols) +
  theme_minimal()+
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  ylab("Proportion of cells")
```

Below the plots:

```{r, fig.width = 6, fig.height = 12}
# Create a grid of the aligned plots
grid <- plot_grid(
  normncells_bar, normncells_bar_B1, normncells_bar_B2,
  nrow = 3,
  align = "v",
  axis = "tb",
  labels = c("A", "B", "C")
)

# Plot ncells per cluster
grid
```

We will normalise internally by replicate (B1 and B2 separately), and after that a second, global normalisation. The numbers are similar but not the same.

```{r}
df_renorm <-
  rbind(df_B1,df_B2)
df_renorm <- df_renorm[df_renorm$CountNorm >0,]

df_renorm$norm <- df_renorm$CountNorm / tapply(df_renorm$CountNorm, df_renorm$Cluster, sum)[df_renorm$Cluster]

# Create a stacked barplot norm number
normncells_bar_renorm <- ggplot(df_renorm, aes(x = Cluster, y = norm, fill = Sample)) +
  geom_col(position = "stack") +
  scale_fill_manual(values = exp_cols) +
  theme_minimal()+
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  ylab("Proportion of cells (normalised by replicate)")

# plot(df_B1$CountNorm[df_B1$CountNorm>0], df_B2$CountNorm[df_B2$CountNorm>0])
# cor(df_B1$CountNorm[df_B1$CountNorm>0], df_B2$CountNorm[df_B2$CountNorm>0])
```


```{r}
normncells_bar_renorm
# pdf("graphics/smed_hnf_barplot_ncell_percluster.pdf", height = 4, width = 6)
# print(normncells_bar_renorm)
# dev.off()
```

## Chi-square analysis

Below a chi-square analysis using a wrapper of the same analysis found in Emili et al., 2025 Science Advances.

```{r, fig.height = 2, fig.width = 8}
library(ComplexHeatmap); library(circlize)
col_enr <- c("#90529f","white","#45aaa7")
col_enr_ramp <- colorRampPalette(col_enr)

source("/omics-storage/lab-solana/alberto/colabos/sizes/code/R_code/r_functions/chisq_and_posthoc.R")

# R1 REPLICATE 1
r1_tbl <- data.frame(table(DF$experiment[DF$replicate=="B1"],DF$cluster[DF$replicate=="B1"])) %>% pivot_wider(names_from=Var1, values_from = Freq) %>%
  column_to_rownames("Var2") %>% as.matrix() %>% as.table
# chi squared
chsq_r1 <- chisq_and_posthoc(r1_tbl)

# matrices
ph_res_1 <- chsq_r1$result[,c(1,2,4)] %>% pivot_wider(names_from = Cluster, values_from = res) %>% column_to_rownames("category") %>% as.matrix()
ph_pva_1 <- chsq_r1$result[,c(1,2,3)] %>% pivot_wider(names_from = Cluster, values_from = pval) %>% column_to_rownames("category") %>% as.matrix()
ph_res_1 <- ph_res_1[,match(levels(DF$cluster),colnames(ph_res_1))]
ph_pva_1 <- ph_pva_1[,match(levels(DF$cluster),colnames(ph_pva_1))]

# stuff for heatmap
col_enr_ramp_1 <- 
  c(
    colorRampPalette(c(col_enr[1],col_enr[2]))(20)[1:19],
    colorRampPalette(c(col_enr[2],col_enr[3]))(21)
    )
enr_pal_broad_1 <- colorRamp2(
  breaks = c(seq(min(ph_res_1),-0.1,length = 19),0,seq(0.1,max(ph_res_1),length = 20)),
  colors = col_enr_ramp(40)
)
pval_fun_broad_1 <- function(j,i,x,y,width,height,fill){
  if(ph_pva_1[i,j] < 0.01){
    if(ph_pva_1[i,j] < 0.001){
      grid.text("**", x, y, gp = gpar(fontsize = 10))
    } else{
      grid.text("*", x, y, gp = gpar(fontsize = 10))
    }
  }
}

ha_chsq_top <- 
  HeatmapAnnotation(
    cluster = 
      anno_simple(
        levels(DF$cluster),
        col = setNames(translate_ids(levels(DF$cluster), smed_hnf_scrna_ctypes_2[,c(1,4)]),levels(DF$cluster)),
        pch = levels(DF$cluster), pt_size = unit(2,"mm")
        ), show_annotation_name = FALSE
    )
ha_chsq_row_r1 <-
  HeatmapAnnotation(
    exp = anno_simple(c("control","HNF4i"), col = setNames(exp_cols[c(1,3)],c("control","HNF4i"))),
    which = "row", show_annotation_name = FALSE
  )
ph_hm_r1 <-
  Heatmap(
    name = "R1",
    ph_res_1,
    col = enr_pal_broad_1,
    cluster_columns = FALSE, cluster_rows = FALSE,
    column_names_side = "top",
    top_annotation = ha_chsq_top,
    column_names_gp = gpar(fontsize = 6),
    row_names_side = "left",
    left_annotation = ha_chsq_row_r1,
    show_column_names = FALSE,
    cell_fun = pval_fun_broad_1
  )

# R2 REPLICATE 2
r2_tbl <- data.frame(table(DF$experiment[DF$replicate=="B2"],DF$cluster[DF$replicate=="B2"])) %>% pivot_wider(names_from=Var1, values_from = Freq) %>%
  column_to_rownames("Var2") %>% as.matrix() %>% as.table
# chi squared
chsq_r2 <- chisq_and_posthoc(r2_tbl)

# matrices
ph_res_2 <- chsq_r2$result[,c(1,2,4)] %>% pivot_wider(names_from = Cluster, values_from = res) %>% column_to_rownames("category") %>% as.matrix()
ph_pva_2 <- chsq_r2$result[,c(1,2,3)] %>% pivot_wider(names_from = Cluster, values_from = pval) %>% column_to_rownames("category") %>% as.matrix()
ph_res_2 <- ph_res_2[,match(levels(DF$cluster),colnames(ph_res_2))]
ph_pva_2 <- ph_pva_2[,match(levels(DF$cluster),colnames(ph_pva_2))]

# stuff for heatmap
col_enr_ramp_2 <- 
  c(
    colorRampPalette(c(col_enr[1],col_enr[2]))(20)[1:19],
    colorRampPalette(c(col_enr[2],col_enr[3]))(21)
  )
enr_pal_broad_2 <- colorRamp2(
  breaks = c(seq(min(ph_res_2),-0.1,length = 19),0,seq(0.1,max(ph_res_2),length = 20)),
  colors = col_enr_ramp(40)
)
pval_fun_broad_2 <- function(j,i,x,y,width,height,fill){
  if(ph_pva_2[i,j] < 0.01){
    if(ph_pva_2[i,j] < 0.001){
      grid.text("**", x, y, gp = gpar(fontsize = 10))
    } else{
      grid.text("*", x, y, gp = gpar(fontsize = 10))
    }
  }
}
ha_chsq_row_r2 <-
  HeatmapAnnotation(
    exp = anno_simple(c("control","HNF4i"), col = setNames(exp_cols[c(1,4)],c("control","HNF4i"))),
    which = "row", show_annotation_name = FALSE
  )
ph_hm_r2 <-
  Heatmap(
    name = "R2",
    ph_res_2,
    col = enr_pal_broad_2,
    cluster_columns = FALSE, cluster_rows = FALSE,
    column_names_side = "top",
    row_names_side = "left",
    left_annotation = ha_chsq_row_r2,
    cell_fun = pval_fun_broad_2
  )

ComplexHeatmap::draw(ph_hm_r1 %v% ph_hm_r2)

# pdf("graphics/smed_hnf_chi2_heatmap.pdf", width = 8, height = 2)
# ComplexHeatmap::draw(ph_hm_r1 %v% ph_hm_r2)
# dev.off()

```

## Correlation

We will quickly check the correlation, pseudobulk-level, of neoblasts, phagocyte progenitors (aberrant or not), and phagocytes, between conditions, to see if and how they are similar to each other.

```{r, fig.height = 6, fig.width = 8}

smed_hnf_scrna$ctype_dge <- smed_hnf_scrna$ctype
smed_hnf_scrna$ctype_dge[smed_hnf_scrna$ctype_dge == "aberrant phagocyte progenitors"] <- "phagocyte progenitors 1"

smed_hnf_counts_ab <- 
  pseudobulk_cond_rep(
    smed_hnf_scrna@assays$RNA@counts,
    identities = smed_hnf_scrna$ctype_dge,
    conditions = smed_hnf_scrna$experiment,
    replicates = rep("REP",length(smed_hnf_scrna$ctype))
  )

m <- smed_hnf_counts_ab$matrix
s <- smed_hnf_counts_ab$sampletable
s$ctype <- factor(s$ctype, levels = levels (smed_hnf_scrna$ctype))
s$condition <- factor(s$condition, levels = c("control","HNF4i"))
s <- s[order(s$ctype, s$condition),]
m <- m[,match(s$sample, colnames(m))]

cl_of_interest <- c("neoblasts","phagocyte progenitors 1", "phagocytes")
m <- m[,colnames(m) %in% s$sample[s$ctype %in% cl_of_interest]]
s <- s[s$ctype %in% cl_of_interest,]
rownames(s) <- gsub("_REP","",s$id_combined)
colnames(m) <- gsub("_REP","",s$id_combined)

hnf_abc_cor_exp <- cor(scale(m))

ha <- HeatmapAnnotation(
  df = s[,c(3,4)],
  col = list(
    ctype = setNames(translate_ids(s$ctype,smed_hnf_scrna_ctypes[,c(2,4)]),s$ctype),
    condition = setNames(c("#efad84","#4a42b5"),c("control","HNF4i"))[match(s$condition, c("control","HNF4i"))]
  ), show_annotation_name = FALSE
)

hb <- HeatmapAnnotation(
  df = s[,c(3,4)],
  col = list(
    ctype = setNames(translate_ids(s$ctype,smed_hnf_scrna_ctypes[,c(2,4)]),s$ctype),
    condition = setNames(c("#efad84","#4a42b5"),c("control","HNF4i"))[match(s$condition, c("control","HNF4i"))]
  ), which = "row", show_legend = FALSE, show_annotation_name = FALSE
)

# custom reordering of the dendrogram
dd.reorder <- 
  reorder(
    as.dendrogram(hclust(dist(hnf_abc_cor_exp, method = "maximum"),method="average")),
    c(5,6,3,4,1,2),
    agglo.FUN = mean
    )

# plot the heatmap
hnf_cors <- Heatmap(
  name = "Pearson",
  hnf_abc_cor_exp,
  cluster_rows = dd.reorder,
  cluster_columns = dd.reorder,
  show_row_names = T,
  show_column_names = T,
  row_names_gp = gpar(fontsize = 8),
  top_annotation = ha,
  left_annotation = hb,
  column_names_gp = gpar(fontsize = 8),
  col = unname(gb_col[c("white1","black1")])#sequential_hcl(10, "YlGnBu", rev = TRUE)
)

ComplexHeatmap::draw(hnf_cors)

# pdf("graphics/smed_hnf_cors.pdf",width = 8, height = 6)
# ComplexHeatmap::draw(hnf_cors)
# dev.off()
```

We see that phagocyte progenitors from the *hnf4* (RNAi) condition are more similar to neoblasts in terms of gene expression.

## Score analysis

Next we will do a gene score analysis.

Gene score is calculated as follows, implementing scanpy's algorithm for gene scoring:

```
gene_score <- function(x, gene_set, gene_pool = rownames(x), remove_set_from_pool = FALSE, fraction = 1, seed = 4343){
  
  # 1. parse gene pool
  # 1.1. remove gene set if asked
  if (remove_set_from_pool == TRUE){
    filt <- !(gene_pool %in% gene_set)
    gene_pool <- gene_pool[filt]
  }
  
  # 1.2. subsample
  n_pool <- length(gene_pool)
  n <- trunc(n_pool*fraction) # if fraction == 1 this will be just all of them
  set.seed(seed) # user can control the seed
  pool <- sample(gene_pool, n)
  
  # 2. calculate score
  
  # genes of interest
  a_i <- apply(x[gene_set,], 2, mean)
  
  # genes from the random pool
  b_i <- apply(x[pool,], 2, mean)
  
  # difference
  score <- a_i - b_i
  
  # 3. result
  return(score)
  
}
```

To choose the gene scores, we will retrieve markers from each cell type.

```{r, fig.height = 4,fig.width = 6}
cl_of_interest_0 <- 
  c("2_neoblasts", "14_aberrant phagocyte progenitors",
    "4_epidermis", "17_phagocyte progenitors 1",
    "11_phagocytes","1_pgrn+ parenchymal cells 2")

smed_cisreg_scrna <- readRDS("outputs/rda/seurat_scrna/smed_cisreg_scRNA_consensus_identities.RDS")
Idents(smed_cisreg_scrna) <- smed_cisreg_scrna$ctype

neoblasts_markers <- 
  FindMarkers(
    smed_cisreg_scrna,
    ident.1 = "neoblasts",
    only.pos = TRUE,
    return.thresh = 1,
    logfc.threshold = 0
  )
markers_neo <- rownames(neoblasts_markers[1:100,])

phagocytes_markers <- 
  FindMarkers(
    smed_cisreg_scrna,
    ident.1 = "phagocytes",
    only.pos = TRUE,
    return.thresh = 1,
    logfc.threshold = 0
  )
markers_phag <- rownames(phagocytes_markers[1:100,])

pgrn_markers <- 
  FindMarkers(
    smed_cisreg_scrna,
    ident.1 = "pgrn+ parenchymal cells 1",
    only.pos = TRUE,
    return.thresh = 1,
    logfc.threshold = 0
  )
markers_pgrn <- rownames(pgrn_markers[1:100,])

epid_markers <- 
  FindMarkers(
    smed_cisreg_scrna,
    ident.1 = "epidermis",
    only.pos = TRUE,
    return.thresh = 1,
    logfc.threshold = 0
  )
markers_epid <- rownames(epid_markers[1:100,])
```

Here we score the cells based on the gene score of each 

```{r}
mat = smed_hnf_scrna@assays$RNA@data
all_genes = rownames(mat)

score_neoblasts <-
  gene_score(
    x = mat, gene_set = markers_neo,
    gene_pool = all_genes, remove_set_from_pool = TRUE,
    fraction = 0.1
  )

score_phagocytes <- 
  gene_score(
    x = mat, gene_set = markers_phag,
    gene_pool = all_genes, remove_set_from_pool = TRUE,
    fraction = 0.1
    )

score_parenchyma <-
  gene_score(
    x = mat, gene_set = markers_pgrn,
    gene_pool = all_genes, remove_set_from_pool = TRUE,
    fraction = 0.1
  )

score_epid <-
  gene_score(
    x = mat, gene_set = markers_epid,
    gene_pool = all_genes, remove_set_from_pool = TRUE,
    fraction = 0.1
  )
```

We make ourselves a data frame where to store all of this. For visualisation purposes, we clip the upper and lower 0.1% values of the distribution of each score.

```{r}
df_scores <-
  data.frame(
    cell = colnames(smed_hnf_scrna), # cells
    exp = smed_hnf_scrna$experiment, # experiment
    rep = smed_hnf_scrna$replicate, # replicates
    exp_rep = smed_hnf_scrna$exp_rep, # merge of replicate and experiment
    clu_maxID = smed_hnf_scrna$clu_maxID, # cluster, name of ctype with max id
    seurat_clusters = smed_hnf_scrna$seurat_clusters, # cluster, name of ctype with max id
    cluster = smed_hnf_scrna$ctype_dge, # cluster
    score_neo = relativise(clip_both(score_neoblasts, Q = 0.999)), # score neoblast markers
    score_pha = relativise(clip_both(score_phagocytes, Q = 0.999)), # score phagocyte markers
    score_par = relativise(clip_both(score_parenchyma, Q = 0.999)), # score parenchyma markers
    score_epi = relativise(clip_both(score_epid, Q = 0.999))   # score epidermis
  )


df_scores <- df_scores[df_scores$clu_maxID %in% cl_of_interest_0,]
df_scores$clu_maxID <- droplevels(df_scores$clu_maxID)
df_scores$cluster <- droplevels(df_scores$cluster)
df_scores <- df_scores %>% mutate(cluster=factor(cluster, labels = c("neoblasts","epidermis","phag.prog","phagocytes","pgrn+ cells")))

# score plot
set.seed(123)
z <- df_scores %>% group_by(exp_rep,cluster) %>% sample_n(50) %>% .$cell
```

Here we create the plots:

```{r}
# Neoblast score plot
boxpl_neo <- ggplot(
  data = df_scores,
  mapping =
    aes(x = cluster,
        y = score_neo,
        fill = exp_rep,
        color = exp_rep)) +
  geom_boxplot(outlier.shape = NA)+
  geom_point(data = df_scores[df_scores$cell %in% z,], mapping = aes(x = cluster, color = exp_rep), shape = 20, position = position_jitterdodge(), size = 1, alpha = .25)+
  theme_classic()+
  scale_fill_manual(values = exp_cols)+
  scale_color_manual(values = darken(exp_cols,0.5))+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size = 8))+
  coord_cartesian(ylim=c(0,1))

# Phagocyte score plot
boxpl_pha <- ggplot(
  data = df_scores,
  mapping =
    aes(x = cluster,
        y = score_pha,
        fill = exp_rep,
        color = exp_rep)) +
  geom_boxplot(outlier.shape = NA)+
  geom_point(data = df_scores[df_scores$cell %in% z,], mapping = aes(x = cluster, color = exp_rep), shape = 20, position = position_jitterdodge(), size = 1, alpha = .25)+
  theme_classic()+
  scale_fill_manual(values = exp_cols)+
  scale_color_manual(values = darken(exp_cols,0.5))+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size = 8))+
  coord_cartesian(ylim=c(0,1))

# Parenchyma score plot
boxpl_par <- ggplot(
  data = df_scores,
  mapping =
    aes(x = cluster,
        y = score_par,
        fill = exp_rep,
        color = exp_rep)) +
  geom_boxplot(outlier.shape = NA)+
  geom_point(data = df_scores[df_scores$cell %in% z,], mapping = aes(x = cluster, color = exp_rep), shape = 20, position = position_jitterdodge(), size = 1, alpha = .25)+
  theme_classic()+
  scale_fill_manual(values = exp_cols)+
  scale_color_manual(values = darken(exp_cols,0.5))+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size = 8))+
  coord_cartesian(ylim=c(0,1))

# Epidermis score plot
boxpl_epi <- ggplot(
  data = df_scores,
  mapping =
    aes(x = cluster,
        y = score_epi,
        fill = exp_rep,
        color = exp_rep)) +
  geom_boxplot(outlier.shape = NA)+
  geom_point(data = df_scores[df_scores$cell %in% z,], mapping = aes(x = cluster, color = exp_rep), shape = 20, position = position_jitterdodge(), size = 1, alpha = .25)+
  theme_classic()+
  scale_fill_manual(values = exp_cols)+
  scale_color_manual(values = darken(exp_cols,0.5))+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size = 8))+
  coord_cartesian(ylim=c(0,1))
```

And here we visualise them:

```{r, fig.width = 15, fig.height = 4}
prow_supp<- plot_grid(
  boxpl_neo + theme(legend.position="none"),
  boxpl_pha + theme(legend.position="none"),
  boxpl_par + theme(legend.position="none"),
  boxpl_epi + theme(legend.position="none"),
  align = 'vh',
  labels = c("Neoblast Score", "Phagocyte Score", "Parenchyma Score", "Epidermis Score"),
  nrow = 1)
plot_grid(prow_supp, lll, rel_widths = c(3, .4))

pdf("graphics/smed_hnf_scores_by_cluster_by_exp_by_rep_SUPP.pdf", height = 4, width = 15)
plot_grid(prow_supp, lll, rel_widths = c(3, .4))
dev.off()
```

Next, a heatmap to show just the median of gene score for the cells from each cluster/condition/experiment. This info is repetitive and does not appear in the manuscript.

```{r, fig.width = 8, fig.height = 2}
test <- df_scores[,-c(1:3,5,6)]
test$group <- paste(test$cluster,test$exp_rep, sep = "__")
test <- aggregate(test[,3:6],by = list(group = test$group), FUN = median) # maybe multiple columns can be passed to aggregate and my life becomes easier
test$exp_rep <- sapply(test$group,function(x){strsplit(x,"__")[[1]][2]})
test$exp_rep <- factor(test$exp_rep, levels = levels(df_scores$exp_rep))
test$cluster <- sapply(test$group,function(x){strsplit(x,"__")[[1]][1]})
test$cluster <- factor(test$cluster, levels = levels(df_scores$cluster))
test <- test[order(test$cluster,test$exp_rep),]

cl_of_interest <- c("neoblasts", "phagocyte progenitors 1","phagocytes","epidermis", "pgrn+ parenchymal cells 2")
col_cls <- setNames(translate_ids(cl_of_interest,smed_ctypes[,c(1,3)]),cl_of_interest)

annot <-
  HeatmapAnnotation(
      cluster = as.character(test$cluster),
      exp_rep = as.character(test$exp_rep),
    col = 
      list(
        cluster = col_cls,
        exp_rep = exp_cols
        )
    )

set.seed(12345678)
Heatmap(
  name = "score",
  t(apply(test[,2:5],2,relativise)),
  cluster_rows = FALSE,
  cluster_columns = TRUE,
  clustering_method_columns = "complete",
  top_annotation = annot,
  col = c("#f4f9ff","#3e4d5d"),
  column_km = 4,
  column_title = NULL,
  show_column_names = FALSE
  )
```

## Significant differences for each score

To see whether and where we have significant differences in the scores shown above, we will use Wilcoxon for each cluster and score.

Since we have two replicates, we opted for doing it for each replicate separately, and for both together (in other words, mingling cells from the same replicates and not regard this). Of note that we will not be showing the plot outputs from each run, only the "both replicates" one. The rest are present in the Manuscript.

First we will declare some functions that are useful, such as boxcox transformation, and a function to remove outliers.

```{r}
library(rcompanion)
library(FSA)

bct <- function(a){
  
  # boxcox model
  bc <- MASS::boxcox(a ~ 1, plotit = FALSE)
  
  # choosing optimal lambda parameter
  optimal <- which( bc$y == max(bc$y) )
  l <- bc$x[optimal]
  
  # transformation
  y <- (a ^ l - 1) / l
}

remove_outliers <- function(x, na.rm = TRUE, ...) {
  qnt <- quantile(x, probs=c(.25, .75), na.rm = na.rm, ...)
  H <- 1.5 * IQR(x, na.rm = na.rm)
  y <- x
  y[x < (qnt[1] - H)] <- NA
  y[x > (qnt[2] + H)] <- NA
  y
}
```

Here's a quick colour-cluster association for the plots that we will see later down below.

```{r}
score_box_cols <- 
  setNames(c("#C8C8C8", "#2171B5", "#80C480", "#228B22", "#FF69B4"), levels(dat$cluster))
```

Here is a big function that will perform the Schreire Ray Hare test of differences (omnibus) followed by a Dunn test and by pairwise comparisons for each cluster. It takes as input the `dat` data frame of gene scores, with boxcox transformed values, the `sco` parameter which indicates what score we are going to compare, and a TRUE/FALSE to remove (or not) the outliers from the individual Wilcoxon tests.

The function creates a formula that will use to model the data for the Scheirer Ray Hare test, followed by post hoc Wilcoxon tests.

```{r}
compare_gene_scores <- function(dat,sco, plot_results = TRUE, remove_outliers_from_individual_wilcox = FALSE){
  f <-
    paste(
      sco, "~", "cluster + exp"
    )
  
  srh <- do.call("scheirerRayHare", list(as.formula(f), data = as.name("dat")))
  
  
  # post hoc tests: propagating cluster and experiment grouping factors to create groups of all combinations
  dat$clu_exp <-
    factor(
      paste(dat$cluster,dat$exp, sep = "|"),
      levels = apply(expand.grid(levels(dat$cluster),levels(dat$exp)), 1, paste, collapse="|")
    )
  
  f2 <- 
    paste(
      sco, "~", "clu_exp"
    )
  
  # dunn test across combinations of clusters and treatments
  DT <-
    do.call(
      "dunnTest",
      list(
        as.formula(f2),
        data = as.name("dat"),
        alpha = 0.005,
        method = "bonferroni"
      )
    )
  
  cld = cldList(P.adj ~ Comparison,
                data = DT,
                threshold = 0.005)
  
  # pairwise wilcoxon test across combinations of clusters and treatments
  pwtest <- pairwise.wilcox.test(
    dat[,sco],
    dat$clu_exp,
    alternative = "two.sided",
    p.adjust.method = "bonferroni"
  )
  
  m_res <- matrix(NA,nrow = length(levels(dat$clu_exp)), ncol = length(levels(dat$clu_exp)))
  diag(m_res) <- 1
  m_res[lower.tri(m_res)] <- pwtest$p.value[complete.cases(pwtest$p.value)]
  m_res[upper.tri(m_res)] <- t(m_res)[upper.tri(m_res)]
  dimnames(m_res) <- list(levels(dat$clu_exp),levels(dat$clu_exp))
  diag(m_res) <- NA
  
  g_ <- 
    graph_from_adjacency_matrix(
      m_res,weighted = TRUE, mode = "undirected")
  
  g_ <- subgraph_from_edges( g_, eids = which(E(g_)$weight < 0.001 ))
  E(g_)$value <- log(-log(E(g_)$weight))
  
  V(g_)$color <- score_box_cols[match(gsub("\\|.*","",V(g_)$name), names(score_box_cols))]
  V(g_)$shape <- "square"
  V(g_)$shape[grep("control",V(g_)$name)] <- "circle"
  V(g_)$label <- "-"
  V(g_)$label[grep("control",V(g_)$name)] <- "+"
  
  colf <- colorRamp2(col = c(alpha("gray",.5),alpha("#241f31",.75)), breaks = c(min(E(g_)$value), max(E(g_)$value)))
  
  E(g_)$col <- colf(E(g_)$value)
  
  set.seed(1234)
  l <- layout_with_graphopt(g_)
  
  
  if(plot_results == TRUE){
    boxplot(
      as.formula(f2),
      dat,
      main = sco,
      col = score_box_cols,
      border = darken(score_box_cols, 0.4),
      outcol = alpha(darken(score_box_cols,.4),.25),
      pch = 20,
      ylim = c(0.4,1)
    )
    plot(g_, edge.color = E(g_)$col, layout = l, edge.width = 3^relativise(E(g_)$value), vertex.label.cex = 2, main = sco)
  }
  
  # individual wilcoxon tests for each cluster
  l_ <- list()
  for(cluster in levels(dat$cluster)){
    d_ <- dat[dat$cluster == cluster,]
    d_$cluster <- droplevels(d_$cluster)
    
    if(remove_outliers_from_individual_wilcox == TRUE){
      d_2 <- d_
      
      d_2[d_2$exp == "control",sco] <- remove_outliers(d_2[d_2$exp == "control",sco])
      d_2[d_2$exp == "HNF4i",sco] <- remove_outliers(d_2[d_2$exp == "HNF4i",sco])
      
      d_ <- d_2
    }
    
    f3 <- paste(sco, "~", "exp")
    w <- 
      do.call(
        "wilcox.test",
        list(
          as.formula(f3),
          data = as.name("d_"),
          alternative = "greater"
        )
      )
    
    l_[[cluster]] <- w
  }
  
  res <-
    list(
      data = dat,
      scheirerRayHare = srh,
      dunnTest = DT,
      dunnTest_comparisons = cld,
      pairwise_wilcox = pwtest,
      pairwise_wilcox_full_matrix = m_res,
      graph = g_,
      graph_layout = l,
      individual_wilcoxon_tests = l_
    )
  
  return(res)
}

# save(df_scores,file = "~/20250224_scores_temporary_backup.rds")
dat <- df_scores
```

First for replicate 1. We subset the data frame of scores, retrieve only cells from Rep B1, and then perform a boxcox transformation of the data. After this, we compare the scores between replicates and conditions:

```{r}
dat_B1 <- dat[dat$rep == "B1",]
dat_B1$exp = factor(dat_B1$exp, levels = c("control", "HNF4i"))

# Box-Cox Transforms, relativised to range between 0-1
dat_B1$score_neo_bc <- relativise(bct(dat_B1$score_neo+0.00000000001))
dat_B1$score_epi_bc <- relativise(bct(dat_B1$score_epi+0.00000000001))
dat_B1$score_pha_bc <- relativise(bct(dat_B1$score_pha+0.00000000001))
dat_B1$score_par_bc <- relativise(bct(dat_B1$score_par+0.00000000001))

# sco <- "score_pha_bc"
neo_sco_comp_B1 <- compare_gene_scores(dat = dat_B1, sco = "score_neo_bc", plot_results = FALSE, remove_outliers_from_individual_wilcox = TRUE)
epi_sco_comp_B1 <- compare_gene_scores(dat = dat_B1, sco = "score_epi_bc", plot_results = FALSE, remove_outliers_from_individual_wilcox = TRUE)
pha_sco_comp_B1 <- compare_gene_scores(dat = dat_B1, sco = "score_pha_bc", plot_results = FALSE, remove_outliers_from_individual_wilcox = TRUE)
par_sco_comp_B1 <- compare_gene_scores(dat = dat_B1, sco = "score_par_bc", plot_results = FALSE, remove_outliers_from_individual_wilcox = TRUE)


pdf("graphics/individual_wilcoxon_scores_B1.pdf", wi = 3.5, he = 9)
par(mfrow = c(4,1))
plot(
  sapply(neo_sco_comp_B1$individual_wilcoxon_tests, function(x){-log(x$p.value)}),
  ylab = "-logp(wilcox.test Control/HNF4i, upper-tail)", main = "neoblasts score", xlab = "",
  type = "b", pch = 21, col = darken(score_box_cols,.4), bg = score_box_cols, cex = 2,
  ylim = c(0,30), bty = "n", axes = FALSE
  )
text(sapply(neo_sco_comp_B1$individual_wilcoxon_tests, function(x){-log(x$p.value)}), labels = round(sapply(neo_sco_comp_B1$individual_wilcoxon_tests, function(x){-log(x$p.value)}),digits = 2),pos = 3)
axis(2)
plot(
  sapply(epi_sco_comp_B1$individual_wilcoxon_tests, function(x){-log(x$p.value)}),
  ylab = "-logp(wilcox.test Control/HNF4i, upper-tail)", main = "epidermis score", xlab = "",
  type = "b", pch = 21, col = darken(score_box_cols,.4), bg = score_box_cols, cex = 2,
  ylim = c(0,30), bty = "n", axes = FALSE
)
text(sapply(epi_sco_comp_B1$individual_wilcoxon_tests, function(x){-log(x$p.value)}), labels = round(sapply(epi_sco_comp_B1$individual_wilcoxon_tests, function(x){-log(x$p.value)}),digits = 2),pos = 3)
axis(2)
plot(
  sapply(pha_sco_comp_B1$individual_wilcoxon_tests, function(x){-log(x$p.value)}),
  ylab = "-logp(wilcox.test Control/HNF4i, upper-tail)", main = "phagocytes score", xlab = "",
  type = "b", pch = 21, col = darken(score_box_cols,.4), bg = score_box_cols, cex = 2,
  ylim = c(0,30), bty = "n", axes = FALSE
)
text(sapply(pha_sco_comp_B1$individual_wilcoxon_tests, function(x){-log(x$p.value)}), labels = round(sapply(pha_sco_comp_B1$individual_wilcoxon_tests, function(x){-log(x$p.value)}),digits = 2),pos = 3)
axis(2)
plot(
  sapply(par_sco_comp_B1$individual_wilcoxon_tests, function(x){-log(x$p.value)}),
  ylab = "-logp(wilcox.test Control/HNF4i, upper-tail)", main = "parenchyma score",
  type = "b", pch = 21, col = darken(score_box_cols,.4), bg = score_box_cols, cex = 2,
  ylim = c(0,100), bty = "n"
)
text(sapply(par_sco_comp_B1$individual_wilcoxon_tests, function(x){-log(x$p.value)}), labels = round(sapply(par_sco_comp_B1$individual_wilcoxon_tests, function(x){-log(x$p.value)}),digits = 2),pos = c(3,3,4,3,3))
par(mfrow = c(1,1))
dev.off()
```

Then for replicate 2. We subset the data frame of scores, retrieve only cells from Rep B2, and then perform a boxcox transformation of the data. After this, we compare the scores between replicates and conditions:

```{r}
dat_B2 <- dat[dat$rep == "B2",]
dat_B2$exp = factor(dat_B2$exp, levels = c("control", "HNF4i"))

# Box-Cox Transforms, relativised to range between 0-1
dat_B2$score_neo_bc <- relativise(bct(dat_B2$score_neo+0.00000000001))
dat_B2$score_epi_bc <- relativise(bct(dat_B2$score_epi+0.00000000001))
dat_B2$score_pha_bc <- relativise(bct(dat_B2$score_pha+0.00000000001))
dat_B2$score_par_bc <- relativise(bct(dat_B2$score_par+0.00000000001))

# sco <- "score_pha_bc"
neo_sco_comp_B2 <- compare_gene_scores(dat = dat_B2, sco = "score_neo_bc", plot_results = FALSE, remove_outliers_from_individual_wilcox = TRUE)
epi_sco_comp_B2 <- compare_gene_scores(dat = dat_B2, sco = "score_epi_bc", plot_results = FALSE, remove_outliers_from_individual_wilcox = TRUE)
pha_sco_comp_B2 <- compare_gene_scores(dat = dat_B2, sco = "score_pha_bc", plot_results = FALSE, remove_outliers_from_individual_wilcox = TRUE)
par_sco_comp_B2 <- compare_gene_scores(dat = dat_B2, sco = "score_par_bc", plot_results = FALSE, remove_outliers_from_individual_wilcox = TRUE)


pdf("graphics/individual_wilcoxon_scores_B2.pdf", wi = 3.5, he = 9)
par(mfrow = c(4,1))
plot(
  sapply(neo_sco_comp_B2$individual_wilcoxon_tests, function(x){-log(x$p.value)}),
  ylab = "-logp(wilcox.test Control/HNF4i, upper-tail)", main = "neoblasts score", xlab = "",
  type = "b", pch = 21, col = darken(score_box_cols,.4), bg = score_box_cols, cex = 2,
  ylim = c(0,30), bty = "n", axes = FALSE
)
text(sapply(neo_sco_comp_B2$individual_wilcoxon_tests, function(x){-log(x$p.value)}), labels = round(sapply(neo_sco_comp_B2$individual_wilcoxon_tests, function(x){-log(x$p.value)}),digits = 2),pos = 3)
axis(2)
plot(
  sapply(epi_sco_comp_B2$individual_wilcoxon_tests, function(x){-log(x$p.value)}),
  ylab = "-logp(wilcox.test Control/HNF4i, upper-tail)", main = "epidermis score", xlab = "",
  type = "b", pch = 21, col = darken(score_box_cols,.4), bg = score_box_cols, cex = 2,
  ylim = c(0,30), bty = "n", axes = FALSE
)
text(sapply(epi_sco_comp_B2$individual_wilcoxon_tests, function(x){-log(x$p.value)}), labels = round(sapply(epi_sco_comp_B2$individual_wilcoxon_tests, function(x){-log(x$p.value)}),digits = 2),pos = 3)
axis(2)
plot(
  sapply(pha_sco_comp_B2$individual_wilcoxon_tests, function(x){-log(x$p.value)}),
  ylab = "-logp(wilcox.test Control/HNF4i, upper-tail)", main = "phagocytes score", xlab = "",
  type = "b", pch = 21, col = darken(score_box_cols,.4), bg = score_box_cols, cex = 2,
  ylim = c(0,30), bty = "n", axes = FALSE
)
text(sapply(pha_sco_comp_B2$individual_wilcoxon_tests, function(x){-log(x$p.value)}), labels = round(sapply(pha_sco_comp_B2$individual_wilcoxon_tests, function(x){-log(x$p.value)}),digits = 2),pos = 3)
axis(2)
plot(
  sapply(par_sco_comp_B2$individual_wilcoxon_tests, function(x){-log(x$p.value)}),
  ylab = "-logp(wilcox.test Control/HNF4i, upper-tail)", main = "parenchyma score",
  type = "b", pch = 21, col = darken(score_box_cols,.4), bg = score_box_cols, cex = 2,
  ylim = c(0,35), bty = "n"
)
text(sapply(par_sco_comp_B2$individual_wilcoxon_tests, function(x){-log(x$p.value)}), labels = round(sapply(par_sco_comp_B2$individual_wilcoxon_tests, function(x){-log(x$p.value)}),digits = 2),pos = c(3,3,4,3,3))
par(mfrow = c(1,1))
dev.off()
```

For both replicates, we just leave it as-is. We show here example plots of what we get from this function:

```{r}
dat_bothreps <- dat
dat_bothreps$exp = factor(dat_bothreps$exp, levels = c("control", "HNF4i"))

# Box-Cox Transforms, relativised to range between 0-1
dat_bothreps$score_neo_bc <- relativise(bct(dat_bothreps$score_neo+0.00000000001))
dat_bothreps$score_epi_bc <- relativise(bct(dat_bothreps$score_epi+0.00000000001))
dat_bothreps$score_pha_bc <- relativise(bct(dat_bothreps$score_pha+0.00000000001))
dat_bothreps$score_par_bc <- relativise(bct(dat_bothreps$score_par+0.00000000001))

# sco <- "score_pha_bc"
neo_sco_comp_bothreps <- compare_gene_scores(dat = dat_bothreps, sco = "score_neo_bc", plot_results = FALSE, remove_outliers_from_individual_wilcox = TRUE)
epi_sco_comp_bothreps <- compare_gene_scores(dat = dat_bothreps, sco = "score_epi_bc", plot_results = FALSE, remove_outliers_from_individual_wilcox = TRUE)
pha_sco_comp_bothreps <- compare_gene_scores(dat = dat_bothreps, sco = "score_pha_bc", plot_results = FALSE, remove_outliers_from_individual_wilcox = TRUE)
par_sco_comp_bothreps <- compare_gene_scores(dat = dat_bothreps, sco = "score_par_bc", plot_results = FALSE, remove_outliers_from_individual_wilcox = TRUE)

par(mfrow = c(4,1))
plot(
  sapply(neo_sco_comp_bothreps$individual_wilcoxon_tests, function(x){-log(x$p.value)}),
  ylab = "-logp(wilcox.test Control/HNF4i, upper-tail)", main = "neoblasts score", xlab = "",
  type = "b", pch = 21, col = darken(score_box_cols,.4), bg = score_box_cols, cex = 2,
  ylim = c(0,30), bty = "n", axes = FALSE
)
text(sapply(neo_sco_comp_bothreps$individual_wilcoxon_tests, function(x){-log(x$p.value)}), labels = round(sapply(neo_sco_comp_bothreps$individual_wilcoxon_tests, function(x){-log(x$p.value)}),digits = 2),pos = 3)
axis(2)
plot(
  sapply(epi_sco_comp_bothreps$individual_wilcoxon_tests, function(x){-log(x$p.value)}),
  ylab = "-logp(wilcox.test Control/HNF4i, upper-tail)", main = "epidermis score", xlab = "",
  type = "b", pch = 21, col = darken(score_box_cols,.4), bg = score_box_cols, cex = 2,
  ylim = c(0,30), bty = "n", axes = FALSE
)
text(sapply(epi_sco_comp_bothreps$individual_wilcoxon_tests, function(x){-log(x$p.value)}), labels = round(sapply(epi_sco_comp_bothreps$individual_wilcoxon_tests, function(x){-log(x$p.value)}),digits = 2),pos = 3)
axis(2)
plot(
  sapply(pha_sco_comp_bothreps$individual_wilcoxon_tests, function(x){-log(x$p.value)}),
  ylab = "-logp(wilcox.test Control/HNF4i, upper-tail)", main = "phagocytes score", xlab = "",
  type = "b", pch = 21, col = darken(score_box_cols,.4), bg = score_box_cols, cex = 2,
  ylim = c(0,45), bty = "n", axes = FALSE
)
text(sapply(pha_sco_comp_bothreps$individual_wilcoxon_tests, function(x){-log(x$p.value)}), labels = round(sapply(pha_sco_comp_bothreps$individual_wilcoxon_tests, function(x){-log(x$p.value)}),digits = 2),pos = 3)
axis(2)
plot(
  sapply(par_sco_comp_bothreps$individual_wilcoxon_tests, function(x){-log(x$p.value)}),
  ylab = "-logp(wilcox.test Control/HNF4i, upper-tail)", main = "parenchyma score",
  type = "b", pch = 21, col = darken(score_box_cols,.4), bg = score_box_cols, cex = 2,
  ylim = c(0,130), bty = "n"
)
text(sapply(par_sco_comp_bothreps$individual_wilcoxon_tests, function(x){-log(x$p.value)}), labels = round(sapply(par_sco_comp_bothreps$individual_wilcoxon_tests, function(x){-log(x$p.value)}),digits = 2),pos = c(3,3,4,3,3))
par(mfrow = c(1,1))

pdf("graphics/individual_wilcoxon_scores_bothreps.pdf", wi = 3.5, he = 9)
par(mfrow = c(4,1))
plot(
  sapply(neo_sco_comp_bothreps$individual_wilcoxon_tests, function(x){-log(x$p.value)}),
  ylab = "-logp(wilcox.test Control/HNF4i, upper-tail)", main = "neoblasts score", xlab = "",
  type = "b", pch = 21, col = darken(score_box_cols,.4), bg = score_box_cols, cex = 2,
  ylim = c(0,30), bty = "n", axes = FALSE
)
text(sapply(neo_sco_comp_bothreps$individual_wilcoxon_tests, function(x){-log(x$p.value)}), labels = round(sapply(neo_sco_comp_bothreps$individual_wilcoxon_tests, function(x){-log(x$p.value)}),digits = 2),pos = 3)
axis(2)
plot(
  sapply(epi_sco_comp_bothreps$individual_wilcoxon_tests, function(x){-log(x$p.value)}),
  ylab = "-logp(wilcox.test Control/HNF4i, upper-tail)", main = "epidermis score", xlab = "",
  type = "b", pch = 21, col = darken(score_box_cols,.4), bg = score_box_cols, cex = 2,
  ylim = c(0,30), bty = "n", axes = FALSE
)
text(sapply(epi_sco_comp_bothreps$individual_wilcoxon_tests, function(x){-log(x$p.value)}), labels = round(sapply(epi_sco_comp_bothreps$individual_wilcoxon_tests, function(x){-log(x$p.value)}),digits = 2),pos = 3)
axis(2)
plot(
  sapply(pha_sco_comp_bothreps$individual_wilcoxon_tests, function(x){-log(x$p.value)}),
  ylab = "-logp(wilcox.test Control/HNF4i, upper-tail)", main = "phagocytes score", xlab = "",
  type = "b", pch = 21, col = darken(score_box_cols,.4), bg = score_box_cols, cex = 2,
  ylim = c(0,45), bty = "n", axes = FALSE
)
text(sapply(pha_sco_comp_bothreps$individual_wilcoxon_tests, function(x){-log(x$p.value)}), labels = round(sapply(pha_sco_comp_bothreps$individual_wilcoxon_tests, function(x){-log(x$p.value)}),digits = 2),pos = 3)
axis(2)
plot(
  sapply(par_sco_comp_bothreps$individual_wilcoxon_tests, function(x){-log(x$p.value)}),
  ylab = "-logp(wilcox.test Control/HNF4i, upper-tail)", main = "parenchyma score",
  type = "b", pch = 21, col = darken(score_box_cols,.4), bg = score_box_cols, cex = 2,
  ylim = c(0,130), bty = "n"
)
text(sapply(par_sco_comp_bothreps$individual_wilcoxon_tests, function(x){-log(x$p.value)}), labels = round(sapply(par_sco_comp_bothreps$individual_wilcoxon_tests, function(x){-log(x$p.value)}),digits = 2),pos = c(3,3,4,3,3))
par(mfrow = c(1,1))
dev.off()
```

## Saving things

```{r}
smed_hnf_scrna_cluster_replicate_identity <-
  unique(data.frame(
    seurat_cluster = smed_hnf_scrna$seurat_clusters,
    ctype = smed_hnf_scrna$ctype,
    broadtype = smed_hnf_scrna$broadtype
  ))

smed_hnf_scrna_cluster_replicate_identity_2 <-
  data.frame(
    seurat_cluster = smed_hnf_scrna$seurat_clusters,
    experiment = smed_hnf_scrna$experiment,
    replicate = smed_hnf_scrna$replicate
  )

smed_hnf_scrna_cluster_replicate_identity_3 <- as.data.frame(table(smed_hnf_scrna_cluster_replicate_identity_2)) %>% mutate(exp_rep = paste(experiment,replicate,sep="_"))

smed_hnf_scrna_cluster_replicate_identity_4 <-
  pivot_wider(
    smed_hnf_scrna_cluster_replicate_identity_3,
    id_cols = seurat_cluster, values_from = Freq,names_from = exp_rep
    )

smed_hnf_scrna_cluster_replicate_identity <- 
  merge(
    smed_hnf_scrna_cluster_replicate_identity,
    smed_hnf_scrna_cluster_replicate_identity_4,
    by.x = 1,
    by.y = 1,
    all.x = TRUE,
    all.y = TRUE
  )

smed_hnf_scrna_cluster_replicate_identity <-
  smed_hnf_scrna_cluster_replicate_identity[
    order(smed_hnf_scrna_cluster_replicate_identity$seurat_cluster),
  ]

xlsx::write.xlsx(
  smed_hnf_scrna_cluster_replicate_identity,
  file = "./outputs/smed_cisreg_hnf_cluster_replicate_identity.xlsx",
  sheetName = "smed hnf scrna cluster identity and replicates",
  col.names = TRUE, row.names = FALSE, showNA = TRUE
)

xlsx::write.xlsx(
  data.frame(
    celltype = c(rep("neoblasts",length(markers_neo)),rep("phagocytes",length(markers_phag))),
    marker = c(markers_neo,markers_phag)
  ),
  file = "./outputs/smed_cisreg_hnf_gene_scores_markers.xlsx",
  sheetName = "smed hnf scrna neoblast and phagocyte markers for score",
  col.names = TRUE, row.names = FALSE, showNA = TRUE
)

xlsx::write.xlsx(
  df_scores,
  file = "./outputs/smed_cisreg_hnf_gene_scores_per_cell.xlsx",
  sheetName = "smed hnf scrna gene scores for cells",
  col.names = TRUE, row.names = FALSE, showNA = TRUE
)

```

## The End!

```{r}
sessionInfo()
```