---
title: "Reanalysis of lhx1/5-1 RNAi single cell DGE"
author: "Alberto Perez-Posada"
date: "2024-12-11"
output: html_document
---

```{r setup, include=FALSE}
dir <- '/omics-storage/lab-solana/alberto/projects/smed_cisreg/'
fcha <- function(){ gsub("-","",Sys.Date()) }

knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = dir)
options(scipen=999)
Sys.setenv(VROOM_CONNECTION_SIZE=5000072)
pdf.options(family = "Helvetica")
```

## About

This markdown reanalyses the knock-down RNAi of lhx1/5-1 in planaria, from Emili et al 2024 ().

We will detect the genes differentially expressed in each cell type and test whether they tend to have a highger predicted activity from the lhx transcription factor.

Part of this code has been adapted from the original publication.

## Load functions

```{r, message = FALSE, warning = FALSE}
library(Matrix)
library(topGO)
library(readr)
library(ggplot2)
library(ggrepel)
library(ggrastr)
```

```{r, message = FALSE, warning = FALSE}
# load functions
source("/omics-storage/lab-solana/alberto/projects/smed_cisreg/code/r_code/functions/sourcefolder.R")
sourceFolder(
  "/omics-storage/lab-solana/alberto/projects/smed_cisreg/code/r_code/functions/", #change for definitive path once its final
  recursive = TRUE
)
```

# Load Data

Important as we will using this gene id quite often here:

```{r}
lhx = "h1SMcG0003793"
```



```{r load_idents}
# Load necessary data
sero_Idents <- read.delim2("/home/alberto/colabos/sero_2024/github_code_orig_upload/planarian-serotonergic-rnai/3_DGE/input/06092023_smed_sero_identities.csv", sep = "," , header = TRUE)
sero_genes <- read.delim2("/home/alberto/colabos/sero_2024/github_code_orig_upload/planarian-serotonergic-rnai/3_DGE/input/06092023_smed_sero_genes.csv", sep = "," , header = TRUE)

sero_Idents[1:5,1:5]
```

Below a table of cell type names, colors etc.

```{r ctypes}
sero_ctypes <- data.frame(
  broad_names = factor(unique(sero_Idents[,c("broad_names")])[c(3,4,11,8,5,2,1,9,6,10,7,12)],levels = unique(sero_Idents[,c("broad_names")])[c(3,4,11,8,5,2,1,9,6,10,7,12)]),
  color = c("#bcbcbc","#0059A9",lighten("royalblue",.3),"limegreen","red3","#ebc054","magenta2","pink","royalblue","#8FCE00","purple","darkgray")
)

sero_ctypes <- merge(
  sero_ctypes,
  as.data.frame(table(sero_Idents$broad_names)),
  by.x = 1,
  by.y = 1,
  all.x = TRUE
)

sero_ctypes<-
  sero_ctypes[order(sero_ctypes$broad_names),]

sero_ctypes
```

We load gene information just in case

```{r rosetta}
rosetta_ext <- read.delim2(
  "/omics-storage/lab-solana/standard_references/Smed_Rink/20231127_Rosetta_EXTENDED.tsv",
  header = TRUE
)
```

## Loading the matrix of gene expression

```{r matrix}
# Load the matrix
sero_X <- readMM(file = "/home/alberto/colabos/sero_2024/github_code_orig_upload/planarian-serotonergic-rnai/3_DGE/input/matrix.mtx.gz")
```

We assign the right gene names and cell names to this matrix

```{r dimnames}
colnames(sero_X) <-  read.table("/home/alberto/colabos/sero_2024/github_code_orig_upload/planarian-serotonergic-rnai/3_DGE/input/barcodes.tsv")[,1]
rownames(sero_X) <-  read.table("/home/alberto/colabos/sero_2024/github_code_orig_upload/planarian-serotonergic-rnai/3_DGE/input/features.tsv")[,1]
```

We subset the matrix to keep the same cells as we had in the original single cell analysis (i.e. those assigned to a cell type.)

```{r subset}
sero_X <- sero_X[,colnames(sero_X) %in% sero_Idents$X]
```

Lastly, because this dataset has multiple interacting terms (biological and technical replicates), we will create what we call "groups". Please see: https://bioconductor.org/packages/devel/bioc/vignettes/DESeq2/inst/doc/DESeq2.html#interactions

```{r rep_group}
rep_group <- data.frame(
  rep = c("B1_T1", "B1_T2", "B2_T1", "B2_T2"),
  group = c("A","B","C","D")
)
```

We bring this information over to the identities metadata table

```{r repgroup_metadata}
sero_Idents$group <- 
  translate_ids(
    x = paste(sero_Idents$biological_rep,sero_Idents$technical_rep, sep = "_"),
    dict = rep_group
  )
```

## Pseudobulk computational dissection/reconstruction of the data

Just like we did with HNF4, we run a special version of our pseudobulk function. We will do this at the broad cell type level as this is the level of resolution of our ANANSE networks.

```{r pseudobulk, message=FALSE}
sero_pseudobulk_cond_reps <- pseudobulk_cond_rep(
  x = sero_X,
  identities = sero_Idents$broad_names,
  conditions = sero_Idents$sample,
  replicates = sero_Idents$group
)
```

We tidy up the table of samples that emerges from this:

```{r cleansampletable}
# Clean sampletable

sero_sampletable <- sero_pseudobulk_cond_reps$sampletable

sero_sampletable_lhx = sero_sampletable[sero_sampletable$condition %in% c("gfp(RNAi)", "lhx1/5-1(RNAi)"), ] # Change based on your conditions
sero_sampletable_lhx <- clean_sampletable(sero_sampletable_lhx)

sero_matrix_lhx <- 
  sero_pseudobulk_cond_reps$matrix[
    ,
    colnames(sero_pseudobulk_cond_reps$matrix) %in% sero_sampletable_lhx$sample
  ]
```

## Differential gene expression analysis over each broad cell type

We will do this at the broad cell type level as this is the level of resolution of our ANANSE networks.

```{r deseq_sc, warning = FALSE, message = FALSE}
sero_DGE_all_lhx <- list()

for(i in unique(sero_sampletable_lhx$ctype)){
  
  print(paste0("starting with cell type ",i))
  
  sero_DGE_all_lhx[[i]] <-
    deseq_pseudobulk(
      count_matrix = sero_matrix_lhx,
      samples_info = sero_sampletable_lhx[,-2],
      celltype = i,
      filter_by = "pvalue", p_threshold = 0.05,
      contrast_info = c("condition","lhx1/5-1(RNAi)","gfp(RNAi)"),
      plot_results = FALSE, min_passing_samples = 2,
      min_counts_per_sample = 1,
      keep_dubious = FALSE
    )
  
  print(paste0("done cell type ",i))

  }
```

A quick look at the differential gene expression results:

```{r volcanoes,fig.width = 12, fig.height = 2.5}
y_max <- 15
library(gridExtra)
list_volcanos <- list()
for(celltype in names(sero_DGE_all_lhx)){
  list_volcanos[[celltype]] <- sero_DGE_all_lhx[[celltype]]$res %>%
    ggplot(aes(x = log2FoldChange, y = -log(pvalue), color = factor(pvalue < 0.05))) +
    geom_point_rast(size = .75) +
    scale_color_manual(values = c("FALSE" = unname(gb_col["black1"]), "TRUE" = "#e66056")) +
    ylim(c(0,y_max))+
    guides(color = "none") + 
    theme_classic() +
    theme(panel.border = element_rect(colour = "black", fill=NA, linewidth = 0.75), axis.title.x=element_blank(), axis.title.y=element_blank())+
    ggtitle(label = celltype)
}

do.call("grid.arrange", c(list_volcanos, ncol=length(list_volcanos)))
```

```{r}
pdf("graphics/RNAi_lhx_volcanoes.pdf", wi = 12, he = 2.5)
do.call("grid.arrange", c(list_volcanos, ncol=length(list_volcanos)))
dev.off()
```

Below a meta-plot of which genes have the highest amount of differentially expressed genes detected.

```{r metaplot}
# METAPLOT DATA FRAME
sero_diffreg_lhx <- data.frame(
  num_diff = sapply(
    sero_DGE_all_lhx,
    function(x){
      a = x$diffgenes
      if(is.na(a[1])){
        b = 0
      } else{
        b = length(a)
      }
      return(b)
    }
  )
)

sero_diffreg_lhx <- 
  merge(
    sero_diffreg_lhx,
    sero_ctypes,
    by.x = 0,
    by.y = "broad_names",
    all.x = TRUE
  )
colnames(sero_diffreg_lhx)[1] <- "ctype"
sero_diffreg_lhx <- sero_diffreg_lhx[-c(grep("unannotated",sero_diffreg_lhx$ctype)),]
```



```{r metaplot_plot, fig.width = 4, fig.height = 4}
##META PLOT NUM DIFF GENES AND CLUSTER SIZE
lhx_metaplot <- ggplot(sero_diffreg_lhx, aes(x = log(Freq), y = num_diff, label = ctype, color = color)) +
  geom_point() +
  labs(
    title = "Sero cluster size vs # DiffReg lhx1/5-1(RNAi)",
    x = "cluster size (log n cells)",
    y = "no. diffreg genes"
  ) +
  scale_color_identity()+
  theme_minimal()+
  theme(panel.border = element_rect(colour = "black", fill=NA, linewidth = 0.75))+
  geom_text_repel(cex=3) + labs(title = "Cluster size vs # DiffReg lhx1/5-1(RNAi)")

lhx_metaplot
```

```{r}
pdf("graphics/RNAi_lhx_metaplot.pdf", he = 4, wi = 4)
lhx_metaplot
dev.off()
```


Below the list of differentially expressed genes. considering all of the possible events (in epidermis only, in neurons only, in parenchyma only, or any combination of the three)

```{r diffgenes}
a = sero_DGE_all_lhx$epidermal$diffgenes
b = sero_DGE_all_lhx$neurons$diffgenes
c = sero_DGE_all_lhx$parenchymal$diffgenes

lhx_diffs <-
  list(
    a = a[!(a %in% b) & !(a %in% c)],
    b = b[!(b %in% a) & !(b %in% c)],
    c = c[!(c %in% a) & !(c %in% b)],
    a_b = a[(a %in% b) & !(a %in% c)],
    a_c = a[(a %in% c) & !(a %in% b)],
    b_c = b[(b %in% c) & !(b %in% a)],
    a_b_c = a[(a %in% b) & (a %in% c)]
  )

sapply(lhx_diffs, length)
```

There are barely any overlaps between these networks, for which we will just keep the genes only differentially regulated in each separate cell type.

```{r diffgenes_2}
lhx_diffs <-
  list(
    epi = a[!(a %in% b) & !(a %in% c)],
    neu = b[!(b %in% a) & !(b %in% c)],
    par = c[!(c %in% a) & !(c %in% b)]
  )

# some colors for some later plots
deg_cols <- setNames(c("#c4c4c4", "#FEC44F", "#2171B5", "#D675AD"),c("none", "neu", "epi", "par"))
deg_outl_col <- darken(deg_cols,.7)
deg_outl_col[1] <- alpha(deg_outl_col[1],.2)
deg_outl_col[c(2,3,4)] <- alpha(deg_outl_col[c(2,3,4)],.25)
```

## Assessing interaction in each network

### epidermis

We will load the epidermis network, parse it accordingly (e.g. subsetting only for interactions stemming from lhx), and add the info of the different DEGs.

```{r nw_epidermis}
# epidermis
epi_nw <- data.table::fread("outputs/ananse/outs/network/epidermis.network", header = TRUE)
epi_nw <- as.data.frame(epi_nw)
epi_nw$tf <- sub("—.*","",epi_nw$tf_target)
epi_nw$tg <- sub(".*—","",epi_nw$tf_target)
epi_nw <- epi_nw[,c(7,8,2,4,5,6)]
df_epi_lhx <- epi_nw[epi_nw$tf == lhx, -1]

# DEGs
df_epi_lhx$deg <- "none"
# DEGs epi
df_epi_lhx$deg[df_epi_lhx$tg %in% lhx_diffs$epi] <- "epi"
# DEGs neu
df_epi_lhx$deg[df_epi_lhx$tg %in% lhx_diffs$neu] <- "neu"
# DEGs par
df_epi_lhx$deg[df_epi_lhx$tg %in% lhx_diffs$par] <- "par"

df_epi_lhx$deg <- factor(df_epi_lhx$deg, levels = c("none","neu","epi","par"))

# Re-normalise the prob and weighted binding scores
df_epi_lhx$rank_prob <- relativise(rank(df_epi_lhx$prob, ties.method = "min"))
df_epi_lhx$rank_wb <- relativise(rank(df_epi_lhx$weighted_binding, ties.method = "min"))
```

For the plot, we will use a stripchart of a subsample of 500 random non-DEG genes plus ALL of the D.E.G.s of each cell type. But the medians and quartiles of the boxes are calculated with all.

```{r boxplot_epi, fig.width = 8, fig.height = 5}
set.seed(123)
non_deg_subsample_epi <- sample(df_epi_lhx$tg[df_epi_lhx$deg == "none"], 500)
f_epi <- which(df_epi_lhx$tg %in% c(non_deg_subsample_epi, df_epi_lhx$tg[df_epi_lhx$deg != "none"]))
par(mfrow = c(1,2))
set.seed(1234)
boxplot(
  df_epi_lhx$prob~df_epi_lhx$deg,
  col = deg_cols, border = darken(deg_cols,.7),
  main = "Interaction score\nof lhx predicted targets\n(epidermis)",
  xlab = "DEGs in lhx K.D.",
  ylab = "Interaction score",
  las = 2,
  cex.main = .8,
  frame.plot = FALSE,
  outline = FALSE
)
stripchart(
  df_epi_lhx$prob[f_epi]~df_epi_lhx$deg[f_epi],
  col = deg_outl_col,
  method = "jitter",
  jitter=0.15,
  vertical = TRUE,
  pch = 20,
  cex=0.7,
  add = TRUE
)
set.seed(1234)
boxplot(
  df_epi_lhx$weighted_binding~df_epi_lhx$deg,
  col = deg_cols, border = darken(deg_cols,.7),
  main = "Weighted binding\nof lhx predicted targets\n(epidermis)",
  xlab = "DEGs in lhx K.D.",
  ylab = "Weighted binding",
  las = 2,
  cex.main = .8,
  frame.plot = FALSE,
  outline = FALSE
)
stripchart(
  df_epi_lhx$weighted_binding[f_epi]~df_epi_lhx$deg[f_epi],
  col = deg_outl_col,
  method = "jitter",
  jitter=0.15,
  vertical = TRUE,
  pch = 20,
  cex=0.7,
  add = TRUE
)
par(mfrow = c(1,1))
```

```{r, echo = FALSE, message = FALSE, warning = FALSE}
pdf("graphics/lhx_rnai_boxplot_ananse_epidermis.pdf",wi = 8, he = 5)
par(mfrow = c(1,2))
set.seed(1234)
boxplot(
  df_epi_lhx$prob~df_epi_lhx$deg,
  col = deg_cols, border = darken(deg_cols,.7),
  main = "Interaction score\nof lhx predicted targets\n(epidermis)",
  xlab = "DEGs in lhx K.D.",
  ylab = "Interaction score",
  las = 2,
  cex.main = .8,
  frame.plot = FALSE,
  outline = FALSE
)
stripchart(
  df_epi_lhx$prob[f_epi]~df_epi_lhx$deg[f_epi],
  col = deg_outl_col,
  method = "jitter",
  jitter=0.15,
  vertical = TRUE,
  pch = 20,
  cex=0.7,
  add = TRUE
)
set.seed(1234)
boxplot(
  df_epi_lhx$weighted_binding~df_epi_lhx$deg,
  col = deg_cols, border = darken(deg_cols,.7),
  main = "Weighted binding\nof lhx predicted targets\n(epidermis)",
  xlab = "DEGs in lhx K.D.",
  ylab = "Weighted binding",
  las = 2,
  cex.main = .8,
  frame.plot = FALSE,
  outline = FALSE
)
stripchart(
  df_epi_lhx$weighted_binding[f_epi]~df_epi_lhx$deg[f_epi],
  col = deg_outl_col,
  method = "jitter",
  jitter=0.15,
  vertical = TRUE,
  pch = 20,
  cex=0.7,
  add = TRUE
)
par(mfrow = c(1,1))
dev.off()
```

### neurons

Loading, parsing and DEG assigning as above.

```{r nw_neurons}
# neuron
neu_nw <- data.table::fread("outputs/ananse/outs/network/neuron.network", header = TRUE)
neu_nw <- as.data.frame(neu_nw)
neu_nw$tf <- sub("—.*","",neu_nw$tf_target)
neu_nw$tg <- sub(".*—","",neu_nw$tf_target)
neu_nw <- neu_nw[,c(7,8,2,4,5,6)]
df_neu_lhx <- neu_nw[neu_nw$tf == lhx, -1]

# neu
df_neu_lhx$deg <- "none"
# DEGs epi
df_neu_lhx$deg[df_neu_lhx$tg %in% lhx_diffs$epi] <- "epi"
# DEGs neu
df_neu_lhx$deg[df_neu_lhx$tg %in% lhx_diffs$neu] <- "neu"
# DEGs par
df_neu_lhx$deg[df_neu_lhx$tg %in% lhx_diffs$par] <- "par"
df_neu_lhx$deg <- factor(df_neu_lhx$deg, levels = c("none","neu","epi","par"))
df_neu_lhx$rank_prob <- relativise(rank(df_neu_lhx$prob, ties.method = "min"))
df_neu_lhx$rank_wb <- relativise(rank(df_neu_lhx$weighted_binding, ties.method = "min"))
```

The plot. The same about the stripchart of the epidermis boxplot above applies here.

```{r boxplot_neu, fig.width = 8, fig.height = 5}
set.seed(123)
non_deg_subsample_neu <- sample(df_neu_lhx$tg[df_neu_lhx$deg == "none"], 500)
f_neu <- which(df_neu_lhx$tg %in% c(non_deg_subsample_neu, df_neu_lhx$tg[df_neu_lhx$deg != "none"]))
par(mfrow = c(1,2))
set.seed(1234)
boxplot(
  df_neu_lhx$prob~df_neu_lhx$deg,
  col = deg_cols, border = darken(deg_cols,.7),
  main = "Interaction score\nof lhx predicted targets\n(neurons)",
  xlab = "DEGs in lhx K.D.",
  ylab = "Interaction score",
  las = 2,
  cex.main = .8,
  frame.plot = FALSE,
  outline = FALSE
)
stripchart(
  df_neu_lhx$prob[f_neu]~df_neu_lhx$deg[f_neu],
  col = deg_outl_col,
  method = "jitter",
  jitter=0.15,
  vertical = TRUE,
  pch = 20,
  cex=0.7,
  add = TRUE
)
set.seed(1234)
boxplot(
  df_neu_lhx$weighted_binding~df_neu_lhx$deg,
  col = deg_cols, border = darken(deg_cols,.7),
  main = "Weighted binding\nof lhx predicted targets\n(neurons)",
  xlab = "DEGs in lhx K.D.",
  ylab = "Weighted binding",
  las = 2,
  cex.main = .8,
  frame.plot = FALSE,
  outline = FALSE
)
stripchart(
  df_neu_lhx$weighted_binding[f_neu]~df_neu_lhx$deg[f_neu],
  col = deg_outl_col,
  method = "jitter",
  jitter=0.15,
  vertical = TRUE,
  pch = 20,
  cex=0.7,
  add = TRUE
)
par(mfrow = c(1,1))
```

```{r}
pdf("graphics/lhx_rnai_boxplot_ananse_neurons.pdf",wi = 8, he = 5)
set.seed(123)
non_deg_subsample_neu <- sample(df_neu_lhx$tg[df_neu_lhx$deg == "none"], 500)
f_neu <- which(df_neu_lhx$tg %in% c(non_deg_subsample_neu, df_neu_lhx$tg[df_neu_lhx$deg != "none"]))

par(mfrow = c(1,2))
set.seed(1234)
boxplot(
  df_neu_lhx$prob~df_neu_lhx$deg,
  col = deg_cols, border = darken(deg_cols,.7),
  main = "Interaction score\nof lhx predicted targets\n(neurons)",
  xlab = "DEGs in lhx K.D.",
  ylab = "Interaction score",
  las = 2,
  cex.main = .8,
  frame.plot = FALSE,
  outline = FALSE
)
stripchart(
  df_neu_lhx$prob[f_neu]~df_neu_lhx$deg[f_neu],
  col = deg_outl_col,
  method = "jitter",
  jitter=0.15,
  vertical = TRUE,
  pch = 20,
  cex=0.7,
  add = TRUE
)
set.seed(1234)
boxplot(
  df_neu_lhx$weighted_binding~df_neu_lhx$deg,
  col = deg_cols, border = darken(deg_cols,.7),
  main = "Weighted binding\nof lhx predicted targets\n(neurons)",
  xlab = "DEGs in lhx K.D.",
  ylab = "Weighted binding",
  las = 2,
  cex.main = .8,
  frame.plot = FALSE,
  outline = FALSE
)
stripchart(
  df_neu_lhx$weighted_binding[f_neu]~df_neu_lhx$deg[f_neu],
  col = deg_outl_col,
  method = "jitter",
  jitter=0.15,
  vertical = TRUE,
  pch = 20,
  cex=0.7,
  add = TRUE
)
par(mfrow = c(1,1))
dev.off()
```

### parenchyma

Loading, parsing and DEG assigning as above.

```{r nw_parenchyma}
# parenchyma
par_nw <- data.table::fread("outputs/ananse/outs/network/parenchyma.network", header = TRUE)
par_nw <- as.data.frame(par_nw)
par_nw$tf <- sub("—.*","",par_nw$tf_target)
par_nw$tg <- sub(".*—","",par_nw$tf_target)
par_nw <- par_nw[,c(7,8,2,4,5,6)]
df_par_lhx <- par_nw[par_nw$tf == lhx, -1]

# par
df_par_lhx$deg <- "none"
# DEGs epi
df_par_lhx$deg[df_par_lhx$tg %in% lhx_diffs$epi] <- "epi"
# DEGs neu
df_par_lhx$deg[df_par_lhx$tg %in% lhx_diffs$neu] <- "neu"
# DEGs par
df_par_lhx$deg[df_par_lhx$tg %in% lhx_diffs$par] <- "par"
df_par_lhx$deg <- factor(df_par_lhx$deg, levels = c("none","neu","epi","par"))
df_par_lhx$rank_prob <- relativise(rank(df_par_lhx$prob, ties.method = "min"))
df_par_lhx$rank_wb <- relativise(rank(df_par_lhx$weighted_binding, ties.method = "min"))

```

The plot. The same about the stripchart of the epidermis boxplot above applies here.

```{r boxplot_parenchyma, fig.width = 8, fig.height = 5}
set.seed(123)
non_deg_subsample_par <- sample(df_par_lhx$tg[df_par_lhx$deg == "none"], 500)
f_par <- which(df_par_lhx$tg %in% c(non_deg_subsample_par, df_par_lhx$tg[df_par_lhx$deg != "none"]))
par(mfrow = c(1,2))
set.seed(1234)
boxplot(
  df_par_lhx$prob~df_par_lhx$deg,
  col = deg_cols, border = darken(deg_cols,.7),
  main = "Interaction score\nof lhx predicted targets\n(parenchyma)",
  xlab = "DEGs in lhx K.D.",
  ylab = "Interaction score",
  las = 2,
  cex.main = .8,
  frame.plot = FALSE,
  outline = FALSE
)
stripchart(
  df_par_lhx$prob[f_par]~df_par_lhx$deg[f_par],
  col = deg_outl_col,
  method = "jitter",
  jitter=0.15,
  vertical = TRUE,
  pch = 20,
  cex=0.7,
  add = TRUE
)
set.seed(1234)
boxplot(
  df_par_lhx$weighted_binding~df_par_lhx$deg,
  col = deg_cols, border = darken(deg_cols,.7),
  main = "Weighted binding\nof lhx predicted targets\n(parenchyma)",
  xlab = "DEGs in lhx K.D.",
  ylab = "Weighted binding",
  las = 2,
  cex.main = .8,
  frame.plot = FALSE,
  outline = FALSE
)
stripchart(
  df_par_lhx$weighted_binding[f_par]~df_par_lhx$deg[f_par],
  col = deg_outl_col,
  method = "jitter",
  jitter=0.15,
  vertical = TRUE,
  pch = 20,
  cex=0.7,
  add = TRUE
)
par(mfrow = c(1,1))
```


```{r}
pdf("graphics/lhx_rnai_boxplot_ananse_parenchyma.pdf",wi = 8, he = 5)
par(mfrow = c(1,2))
set.seed(1234)
boxplot(
  df_par_lhx$prob~df_par_lhx$deg,
  col = deg_cols, border = darken(deg_cols,.7),
  main = "Interaction score\nof lhx predicted targets\n(parenchyma)",
  xlab = "DEGs in lhx K.D.",
  ylab = "Interaction score",
  las = 2,
  cex.main = .8,
  frame.plot = FALSE,
  outline = FALSE
)
stripchart(
  df_par_lhx$prob[f_par]~df_par_lhx$deg[f_par],
  col = deg_outl_col,
  method = "jitter",
  jitter=0.15,
  vertical = TRUE,
  pch = 20,
  cex=0.7,
  add = TRUE
)
set.seed(1234)
boxplot(
  df_par_lhx$weighted_binding~df_par_lhx$deg,
  col = deg_cols, border = darken(deg_cols,.7),
  main = "Weighted binding\nof lhx predicted targets\n(parenchyma)",
  xlab = "DEGs in lhx K.D.",
  ylab = "Weighted binding",
  las = 2,
  cex.main = .8,
  frame.plot = FALSE,
  outline = FALSE
)
stripchart(
  df_par_lhx$weighted_binding[f_par]~df_par_lhx$deg[f_par],
  col = deg_outl_col,
  method = "jitter",
  jitter=0.15,
  vertical = TRUE,
  pch = 20,
  cex=0.7,
  add = TRUE
)
par(mfrow = c(1,1))
dev.off()
```


## The End!

```{r}
sessionInfo()
```
