---
title: "Reanalysis of alx3-1 RNAi (Akheralie et al) and leverage to TF/target networks"
author: "Alberto Perez-Posada"
date: "2024-12-11"
output: html_document
---

```{r setup, include=FALSE}
dir <- '/omics-storage/lab-solana/alberto/projects/smed_cisreg/'
fcha <- function(){ gsub("-","",Sys.Date()) }

knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = dir)
options(scipen=999)
Sys.setenv(VROOM_CONNECTION_SIZE=5000072)
pdf.options(family = "Helvetica")
```

## About

This markdown reanalyses the knock-down RNAi of alx3-1 in planaria, from Akheralie et al., 2023 (PMID: 37681295). For more information please visit: https://doi.org/10.1242/dev.201777.

Here we used bowtie2 for the mapping of reads, quantification with HTSeq, and differential gene expression with DESeq2.

We will retrieve a set of differentially expressed genes as done in the original manuscript, and bring them over to the networks of the different cell types to test whether they tend to have a higher predicted activity by coe.

## Load libraries and functions

We will load the functions as we usually do:

```{r, message = FALSE, warning = FALSE}
library(countToFPKM)
library(ComplexHeatmap)
library(Mfuzz)
library(dplyr)
library(RColorBrewer)
library(colorspace)
library(viridis)
library(Matrix)
library(topGO)
library(readr)
library(ggplot2)
library(ggrepel)
library(ggrastr)
library(DESeq2)

# load functions
source("/omics-storage/lab-solana/alberto/projects/smed_cisreg/code/r_code/functions/sourcefolder.R")
sourceFolder(
  "/omics-storage/lab-solana/alberto/projects/smed_cisreg/code/r_code/functions/", #change for definitive path once its final
  recursive = TRUE
)
```

## Load Data

We will load a bunch of data at the beginning of the markdown. Leveraging gene expression and gene network data: despite this (and other) knock-down assays were performed using bulk RNA-Seq in whole animals, we can integrate this information with our network data and ask the question of whether such genes have any different predicted interactions. If that was the case, one premise is that we should be able detect which are the cell types in which we can expect to find such genes. For that we can interrogate our single cell data. Are there any cell types expressing them more often?

We will load here the single cell data to explore this later.

```{r}
scdata <- readRDS("outputs/rda/seurat_scrna/smed_cisreg_scRNA_consensus_identities.RDS")

scdata <- readRDS("outputs/rda/seurat_scrna/smed_cisreg_scRNA_consensus_identities.RDS")
broadcols <-
  setNames(
    c(
      "#969696", # neoblast
      "#9ECAE1", # EEP
      "#56A0CE", # EEP
      "#2171B5", # epidermis
      "#228B22", # phagocytes
      "#869E42", # basal/goblet
      "#B22222", # muscle
      "#FEC44F", # neurons
      "#FF69B4", # parenchyma
      "#FFB0AB", # protonephridia
      "#6428C8"  # secretory
    ),
    unique(sort(scdata$broadtype_splitbam))[c(6,2,4,3,9,1,5,7,8,10,11)]
  )

mat = scdata@assays$RNA@data
all_genes = rownames(mat)
```

We will load the pseudobulk counts to quantify the max expression of every gene. From this we can get the genes with a max of expresison in muscle, neurons, and secretory cells.

```{r}
load("outputs/rda/smed_counts.rda")
# a vector of which is the cell type with highest expression for each gene (using our pseudobulk data).
smed_maxc <- setNames(apply(smed_counts_norm_cw, 1, function(x){names(x[x==max(x)])}),rownames(smed_counts_norm_cw))

maxc_sec <- names(smed_maxc[grep("secretory",smed_maxc)])
maxc_neu <- names(smed_maxc[smed_maxc %in% grep("neuron", colnames(smed_counts), va = T) ])
maxc_mus <- names(smed_maxc[smed_maxc %in% grep("muscle", colnames(smed_counts), va = T) ])
```

Then we will load the network data of muscle, neurons, and secretory cells.

```{r}
# Load networks, keep them there
mus_nw <- data.table::fread("outputs/ananse/outs/network/muscle.network", header = TRUE)
mus_nw <- as.data.frame(mus_nw)
mus_nw$tf <- sub("—.*","",mus_nw$tf_target)
mus_nw$tg <- sub(".*—","",mus_nw$tf_target)
mus_nw <- mus_nw[,c(7,8,2,4,5,6)]

neu_nw <- data.table::fread("outputs/ananse/outs/network/neuron.network", header = TRUE)
neu_nw <- as.data.frame(neu_nw)
neu_nw$tf <- sub("—.*","",neu_nw$tf_target)
neu_nw$tg <- sub(".*—","",neu_nw$tf_target)
neu_nw <- neu_nw[,c(7,8,2,4,5,6)]

sec_nw <- data.table::fread("outputs/ananse/outs/network/secretory.network", header = TRUE)
sec_nw <- as.data.frame(sec_nw)
sec_nw$tf <- sub("—.*","",sec_nw$tf_target)
sec_nw$tg <- sub(".*—","",sec_nw$tf_target)
sec_nw <- sec_nw[,c(7,8,2,4,5,6)]
```

Let's keep the gene ID of alx3-1 at hand:

```{r}
alx <- "h1SMcG0021040"
```

```{r}
p_alx <-
  FeaturePlot(
    scdata,
    reduction = "umap",
    dims = c(1,2),
    features = alx,
    cols = colorRampPalette(c("lightgrey","#a100d2","#4c1f89"))(10),
    pt.size = 0.5,
    order = T,
    label = FALSE,
    slot = "counts"
  )+NoAxes()+NoLegend()

png(paste0("graphics/umap_alx_",alx,".png"), wi = 256, he = 256)
print(p_alx)
dev.off()

umap_broad <-
  DimPlot(
    scdata,
    reduction = "umap",
    dims = c(1,2),
    group.by = "broadtype",
    cols = setNames(broadcols, sub("_go","/go", names(broadcols))),
    pt.size = 0.5,
    label = FALSE
  )+NoAxes()+NoLegend()

png(paste0("graphics/umap_broad.png"), wi = 256, he = 256)
print(umap_broad)
dev.off()
```

## Differential Gene Expression

Here we load the sample information (which is basically the SRA metadata table)

```{r}
# Load files information
samplesdir <- "/home/alberto/storage/projects/smed_cisreg/outputs/rna/alx3_akheralie/intact/salmon_output/"
files <- list.files(path = samplesdir,pattern = ".sf",recursive = TRUE)

sampletable <- 
  read.delim2(
    "/home/alberto/storage/projects/smed_cisreg/outputs/rna/alx3_akheralie/intact/sampletable.tsv",
    header = TRUE
  )
rownames(sampletable) <- sampletable$Run
sampletable$condition <- factor(sampletable$condition, levels = c("control(RNAi)","alx-3(RNAi)"))
sampletable$batch <- factor(sampletable$batch, levels = c(1,2,3))
sampletable <- sampletable[,c("Run","condition","batch")]

head(sampletable)
```

We load the samples here using `tximport`:

```{r}
# Load the counts
txi <- tximport(paste0(samplesdir,files), type = "salmon", tx2gene = rosetta_ext[,c(2,1)])
```

## Differential Gene Expression

We create a DESeq object from the count data and the sample table, using the design batch + condition.

```{r}
# DESeq dds object
dds <- DESeqDataSetFromTximport(txi,colData = sampletable, ~batch+condition)
```

We will have a quick look at the replicates by retrieving the variance-stabilised expression data from the object, clustering, and plotting a PCA.

```{r}
vsd <- vst(dds, blind=FALSE)
plotPCA(vsd, ntop = 200, intgroup = c("condition","batch")) # we can consider removing replicate 2 from the assay tbh
```

Also a heatmap:

```{r, fig.width = 4.2, fig.height = 3.5}
sampleDists <- dist(t(assay(vsd)))
h <- hclust(sampleDists)
h_alx_samples <-
  Heatmap(
    name = "Pearson",
    cor(log1p(assay(vsd))),
    cluster_rows = h,
    cluster_columns = h,
    col = sequential_hcl(10,"YlOrRd", rev = T),
    top_annotation = 
      HeatmapAnnotation(
        batch = sampletable$batch,
        condition = sampletable$condition,
        col = 
          list(
            batch = setNames(colorspace::lighten(c("red","green","blue"),.2), c("1","2","3")),
            condition = setNames(c("orange","purple"), c("control(RNAi)","alx-3(RNAi)"))
          )
      ),
    left_annotation = 
      HeatmapAnnotation(
        batch = sampletable$batch,
        condition = sampletable$condition,
        col = 
          list(
            batch = setNames(colorspace::lighten(c("red","green","blue"),.2), c("1","2","3")),
            condition = setNames(c("orange","purple"), c("control(RNAi)","alx-3(RNAi)"))
          ),
        which = "row",
        show_legend = FALSE
      )
  )

draw(h_alx_samples)

pdf("graphics/alx_heatmap_samples.pdf", wi = 5.5, he = 4)
draw(h_alx_samples)
dev.off()
```

We follow with estimation of size factors and dispersions, and the Wald test.

```{r}
dds <- estimateSizeFactors(dds)
dds <- estimateDispersions(dds)
dds <- nbinomWaldTest(dds)
```

```{r}
res <- results(dds, contrast=c("condition","alx-3(RNAi)","control(RNAi)"))
res <- res[complete.cases(res),]

fc_thres <- log(1.5,2) # a foldchange of 1.5x
p_thres <- 0.1

f_signif <- res$padj < p_thres & abs(res$log2FoldChange) > fc_thres

res$deg <- ifelse(f_signif, "DEG", "none")
```

The volcano plot:

```{r, fig.width = 3.5, fig.height = 3.5}
v_alx <-
  volcano_plot(res,fc_thres = fc_thres, p_thres = p_thres, main = "Diff. Exp. Genes in Akheralie et al., 2023", ylim = 120)

print(v_alx)


```

```{r}
pdf("graphics/alx_volcano.pdf", wi = 4, he = 5)
print(v_alx)
dev.off()
```

Because we are only interested in down-regulated, we will subset them.

```{r}
alx_diffreg <- rownames(res)[f_signif]
alx_down <- rownames(res)[res$padj < p_thres & res$log2FoldChange < - fc_thres ]
```

```{r, fig.width = 5, fig.height = 6}
h_alx <- 
  Heatmap(
    name = "expression",
    t(scale(t(assay(vsd[rownames(vsd) %in% alx_diffreg,])))),
    clustering_method_rows = "ward.D2",
    clustering_method_columns = "ward.D2",
    col = rev(brewer.pal(10,"RdYlBu")),
    show_row_names = FALSE,
    show_column_names = FALSE,
    right_annotation = 
      HeatmapAnnotation(
        diffreg = setNames(
          ifelse(
            rownames(vsd[rownames(vsd) %in% alx_diffreg,]) %in% alx_diffreg & rownames(vsd[rownames(vsd) %in% alx_diffreg,]) %in% alx_down,
            "DOWN",
            "UP"
          ),
          rownames(vsd[rownames(vsd) %in% alx_diffreg,])
        ),
        col = list(diffreg = setNames(c("lightgreen","tomato"),c("UP","DOWN"))),
        which = "row"
      ),
    top_annotation = 
      HeatmapAnnotation(
        batch = sampletable$batch,
        condition = sampletable$condition,
        col = 
          list(
            batch = setNames(colorspace::lighten(c("red","green","blue"),.2), c("1","2","3")),
            condition = setNames(c("orange","purple"), c("control(RNAi)","alx-3(RNAi)"))
          )
      )
  )

pdf("graphics/alx_heatmap.pdf", wi = 5, he = 6)
draw(h_alx)
dev.off()

draw(h_alx)
```

## Gene score of alx3-1

To assess which cell types are expressing our genes more, we can calculate a "gene score" as done in other markdowns.

```{r}
score_alx <-
  gene_score(
    x = mat, gene_set = alx_down[alx_down %in% rownames(mat)],
    gene_pool = all_genes, remove_set_from_pool = TRUE,
    fraction = 0.1
  )

```

We will create a data frame for a later boxplot.

To accomodate for outliers in the distributions, we will clip the data from outliers (bottom 5% and top 5%).

```{r}
df_scores_alx <-
  data.frame(
    cell = colnames(scdata), # cells
    clu_narrow = scdata$ctype,
    cluster = factor(scdata$broadtype_splitbam, levels = unique(sort(scdata$broadtype_splitbam))[c(6,2,4,3,9,1,5,7,8,10,11)]), # cluster
    score_alx = relativise(clip_q(clip_q(score_alx,q=.05,method = "down"), method = "up", q = .95))
  )
```

We create the plot here:

```{r}
# score plot
set.seed(123)
z <- df_scores_alx %>% group_by(cluster) %>% sample_n(200) %>% .$cell
df_scores_alx$jit <- ifelse(df_scores_alx$cell %in% z, TRUE, FALSE)

# score plot
boxpl_score_alx <-
  df_scores_alx %>%
  ggplot(mapping = aes(x = cluster, y = score_alx, fill = cluster, color = cluster)) +
  geom_boxplot(outlier.shape = NA)+
  geom_jitter(mapping = aes(shape=jit),size = .6, alpha = .25)+
  theme_classic()+
  scale_fill_manual(values = broadcols)+
  scale_color_manual(values = darken(broadcols,0.5))+
  scale_shape_manual(values = setNames(c(20,NA),c("TRUE","FALSE")))+
  theme(axis.text.x = element_blank(), legend.position = "none")+
  ylab(label = "score alx3-1 (RNAi) downregulated genes")+
  coord_cartesian(ylim=c(0,1))
```

We plot it here:

```{r, fig.height = 3.5, fig.width = 3.5}
boxpl_score_alx
```

```{r}
pdf("graphics/alx_downreg_genescore.pdf", he = 3.5, wi = 3.5)
boxpl_score_alx
dev.off()
```

We confirm indeed the potential targets of prep are mostly expressed in the same cell types where we detected coe as influential (the epidermal lineage and neurons).

## Leveraging differential expression and network data

We will leverage these predicted ANANSE interaction scores with the gene expression data, to check if prep downregulated genes have a higher interaction in our networks.

Provided we have single cell data to do the following, we will also incorporate a subfiltering to keep only the D.E.G.s whose higher expression is in the broad cell type of choice (muscle, neurons, secretory). That is, prep RNAi D.E.G.s that are normally detected as expressed in muscle, neurons, or secretory cells.

We incorporate this in a column of the data frame. We will also create a filter of 500 random non-D.E.G.s in order to do a stripchart on top of the boxplot.

### Muscle

Regarding muscle:

```{r}
# muscle
df_mus_alx <- mus_nw[mus_nw$tf == alx, -1]

# DEGs alx
df_mus_alx$deg <- 
  factor(ifelse(
    df_mus_alx$tg %in% alx_down & 
      df_mus_alx$tg %in% maxc_mus
    , "down","none"), levels = c("none","down"))

df_mus_alx$rank_prob <- relativise(rank(df_mus_alx$prob, ties.method = "min"))
df_mus_alx$rank_wb <- relativise(rank(df_mus_alx$weighted_binding, ties.method = "min"))

head(df_mus_alx)

set.seed(123)
non_deg_subsample_mus <- sample(df_mus_alx$tg[df_mus_alx$deg == "none"], 500)

f_mus <- which(df_mus_alx$tg %in% c(non_deg_subsample_mus,df_mus_alx$tg[df_mus_alx$deg != "none"]))

deg_cols <- setNames(c("#c4c4c4","#B22222"),c("none","down"))
deg_outl_col <- deg_cols
deg_outl_col <- darken(deg_outl_col,.7)
deg_outl_col[1] <- alpha(deg_outl_col[1],.2)
deg_outl_col[c(2,3,4)] <- alpha(deg_outl_col[c(2,3,4)],.25)
```

And here the plot:

```{r, fig.width = 5, fig.height = 5}
par(mfrow = c(1,2))
set.seed(1234)
boxplot(
  df_mus_alx$prob~df_mus_alx$deg,
  col = deg_cols, border = darken(deg_cols,.7),
  main = "Interaction score\nof alx predicted targets\n(musron)",
  xlab = "DEGs in alx K.D. (Akheralie et al.)",
  ylab = "Interaction score",
  las = 2,
  cex.main = .6,
  frame.plot = FALSE,
  outline = FALSE
)
stripchart(
  df_mus_alx$prob[f_mus]~df_mus_alx$deg[f_mus],
  col = deg_outl_col,
  method = "jitter",
  jitter=0.15,
  vertical = TRUE,
  pch = 20,
  cex=0.7,
  add = TRUE
)
set.seed(1234)
boxplot(
  df_mus_alx$weighted_binding~df_mus_alx$deg,
  col = deg_cols, border = darken(deg_cols,.7),
  main = "Weighted binding\nof alx predicted targets\n(muscle)",
  xlab = "DEGs in alx K.D. (Akheralie et al.)",
  ylab = "Weighted binding",
  las = 2,
  cex.main = .6,
  frame.plot = FALSE,
  outline = FALSE
)
stripchart(
  df_mus_alx$weighted_binding[f_mus]~df_mus_alx$deg[f_mus],
  col = deg_outl_col,
  method = "jitter",
  jitter=0.15,
  vertical = TRUE,
  pch = 20,
  cex=0.7,
  add = TRUE
)
par(mfrow = c(1,1))
```

### Neurons

```{r}
# neuron
df_neu_alx <- neu_nw[neu_nw$tf == alx, -1]

# DEGs alx
df_neu_alx$deg <- 
  factor(ifelse(
    df_neu_alx$tg %in% alx_down & 
      df_neu_alx$tg %in% maxc_neu
    , "down","none"), levels = c("none","down"))

df_neu_alx$rank_prob <- relativise(rank(df_neu_alx$prob, ties.method = "min"))
df_neu_alx$rank_wb <- relativise(rank(df_neu_alx$weighted_binding, ties.method = "min"))

head(df_neu_alx)

set.seed(123)
non_deg_subsample_neu <- sample(df_neu_alx$tg[df_neu_alx$deg == "none"], 500)

f_neu <- which(df_neu_alx$tg %in% c(non_deg_subsample_neu,df_neu_alx$tg[df_neu_alx$deg != "none"]))

deg_cols <- setNames(c("#c4c4c4","#FEC44F"),c("none","down"))
deg_outl_col <- deg_cols
deg_outl_col <- darken(deg_outl_col,.7)
deg_outl_col[1] <- alpha(deg_outl_col[1],.2)
deg_outl_col[c(2,3,4)] <- alpha(deg_outl_col[c(2,3,4)],.25)
```

```{r, fig.width = 5, fig.height = 5}
par(mfrow = c(1,2))
set.seed(1234)
boxplot(
  df_neu_alx$prob~df_neu_alx$deg,
  col = deg_cols, border = darken(deg_cols,.7),
  main = "Interaction score\nof alx predicted targets\n(neuron)",
  xlab = "DEGs in alx K.D. (Akheralie et al.)",
  ylab = "Interaction score",
  las = 2,
  cex.main = .6,
  frame.plot = FALSE,
  outline = FALSE
)
stripchart(
  df_neu_alx$prob[f_neu]~df_neu_alx$deg[f_neu],
  col = deg_outl_col,
  method = "jitter",
  jitter=0.15,
  vertical = TRUE,
  pch = 20,
  cex=0.7,
  add = TRUE
)
set.seed(1234)
boxplot(
  df_neu_alx$weighted_binding~df_neu_alx$deg,
  col = deg_cols, border = darken(deg_cols,.7),
  main = "Weighted binding\nof alx predicted targets\n(neuron)",
  xlab = "DEGs in alx K.D. (Akheralie et al.)",
  ylab = "Weighted binding",
  las = 2,
  cex.main = .6,
  frame.plot = FALSE,
  outline = FALSE
)
stripchart(
  df_neu_alx$weighted_binding[f_neu]~df_neu_alx$deg[f_neu],
  col = deg_outl_col,
  method = "jitter",
  jitter=0.15,
  vertical = TRUE,
  pch = 20,
  cex=0.7,
  add = TRUE
)
par(mfrow = c(1,1))
```

### Secretory

```{r}
# secretory
df_sec_alx <- sec_nw[sec_nw$tf == alx, -1]

# DEGs alx
df_sec_alx$deg <- 
  factor(ifelse(
    df_sec_alx$tg %in% alx_down # &
    # df_sec_alx$tg %in% maxc_sec
    , "down","none"
  ), levels = c("none","down"))

df_sec_alx$rank_prob <- relativise(rank(df_sec_alx$prob, ties.method = "min"))
df_sec_alx$rank_wb <- relativise(rank(df_sec_alx$weighted_binding, ties.method = "min"))

set.seed(123)
non_deg_subsample_sec <- sample(df_sec_alx$tg[df_sec_alx$deg == "none"], 500)

f_sec <- which(df_sec_alx$tg %in% c(non_deg_subsample_sec,df_sec_alx$tg[df_sec_alx$deg != "none"]))
```

```{r}
par(mfrow = c(1,2))
deg_cols <- setNames(c("#c4c4c4","#6428C8"),c("none","down"))
deg_outl_col <- deg_cols
deg_outl_col <- darken(deg_outl_col,.7)
deg_outl_col[1] <- alpha(deg_outl_col[1],.2)
deg_outl_col[c(2,3,4)] <- alpha(deg_outl_col[c(2,3,4)],.25)

set.seed(1234)
boxplot(
  df_sec_alx$prob~df_sec_alx$deg,
  col = deg_cols, border = darken(deg_cols,.7),
  main = "Interaction score\nof alx predicted targets\n(secretory)",
  xlab = "DEGs in alx K.D. (Akheralie et al.)",
  ylab = "Interaction score",
  las = 2,
  cex.main = .6,
  frame.plot = FALSE,
  outline = FALSE
)
stripchart(
  df_sec_alx$prob[f_sec]~df_sec_alx$deg[f_sec],
  col = deg_outl_col,
  method = "jitter",
  jitter=0.15,
  vertical = TRUE,
  pch = 20,
  cex=0.7,
  add = TRUE
)

set.seed(1234)
boxplot(
  df_sec_alx$weighted_binding~df_sec_alx$deg,
  col = deg_cols, border = darken(deg_cols,.7),
  main = "Weighted binding\nof alx predicted targets\n(secretory)",
  xlab = "DEGs in alx K.D. (Akheralie et al.)",
  ylab = "Weighted binding",
  las = 2,
  cex.main = .6,
  frame.plot = FALSE,
  outline = FALSE
)
stripchart(
  df_sec_alx$weighted_binding[f_sec]~df_sec_alx$deg[f_sec],
  col = deg_outl_col,
  method = "jitter",
  jitter=0.15,
  vertical = TRUE,
  pch = 20,
  cex=0.7,
  add = TRUE
)
par(mfrow = c(1,1))
```

### Plot of all together

```{r}
pdf("graphics/alx_ananse_score.pdf",wi = 9, he = 3.5)
par(mfrow = c(1,6))

# MUSCLE
set.seed(123)
non_deg_subsample_mus <- sample(df_mus_alx$tg[df_mus_alx$deg == "none"], 500)

f_mus <- which(df_mus_alx$tg %in% c(non_deg_subsample_mus,df_mus_alx$tg[df_mus_alx$deg != "none"]))

deg_cols <- setNames(c("#c4c4c4","#B22222"),c("none","down"))
deg_outl_col <- deg_cols
deg_outl_col <- darken(deg_outl_col,.7)
deg_outl_col[1] <- alpha(deg_outl_col[1],.2)
deg_outl_col[c(2,3,4)] <- alpha(deg_outl_col[c(2,3,4)],.25)

set.seed(1234)
boxplot(
  df_mus_alx$prob~df_mus_alx$deg,
  col = deg_cols, border = darken(deg_cols,.7),
  main = "Interaction score\nof alx predicted targets\n(musron)",
  xlab = "DEGs in alx K.D. (Akheralie et al.)",
  ylab = "Interaction score",
  las = 2,
  cex.main = .6,
  frame.plot = FALSE,
  outline = FALSE
)
stripchart(
  df_mus_alx$prob[f_mus]~df_mus_alx$deg[f_mus],
  col = deg_outl_col,
  method = "jitter",
  jitter=0.15,
  vertical = TRUE,
  pch = 20,
  cex=0.7,
  add = TRUE
)
set.seed(1234)
boxplot(
  df_mus_alx$weighted_binding~df_mus_alx$deg,
  col = deg_cols, border = darken(deg_cols,.7),
  main = "Weighted binding\nof alx predicted targets\n(muscle)",
  xlab = "DEGs in alx K.D. (Akheralie et al.)",
  ylab = "Weighted binding",
  las = 2,
  cex.main = .6,
  frame.plot = FALSE,
  outline = FALSE
)
stripchart(
  df_mus_alx$weighted_binding[f_mus]~df_mus_alx$deg[f_mus],
  col = deg_outl_col,
  method = "jitter",
  jitter=0.15,
  vertical = TRUE,
  pch = 20,
  cex=0.7,
  add = TRUE
)

# NEURONS
set.seed(123)
non_deg_subsample_neu <- sample(df_neu_alx$tg[df_neu_alx$deg == "none"], 500)

f_neu <- which(df_neu_alx$tg %in% c(non_deg_subsample_neu,df_neu_alx$tg[df_neu_alx$deg != "none"]))

deg_cols <- setNames(c("#c4c4c4","#FEC44F"),c("none","down"))
deg_outl_col <- deg_cols
deg_outl_col <- darken(deg_outl_col,.7)
deg_outl_col[1] <- alpha(deg_outl_col[1],.2)
deg_outl_col[c(2,3,4)] <- alpha(deg_outl_col[c(2,3,4)],.25)

set.seed(1234)
boxplot(
  df_neu_alx$prob~df_neu_alx$deg,
  col = deg_cols, border = darken(deg_cols,.7),
  main = "Interaction score\nof alx predicted targets\n(neuron)",
  xlab = "DEGs in alx K.D. (Akheralie et al.)",
  ylab = "Interaction score",
  las = 2,
  cex.main = .6,
  frame.plot = FALSE,
  outline = FALSE
)
stripchart(
  df_neu_alx$prob[f_neu]~df_neu_alx$deg[f_neu],
  col = deg_outl_col,
  method = "jitter",
  jitter=0.15,
  vertical = TRUE,
  pch = 20,
  cex=0.7,
  add = TRUE
)
set.seed(1234)
boxplot(
  df_neu_alx$weighted_binding~df_neu_alx$deg,
  col = deg_cols, border = darken(deg_cols,.7),
  main = "Weighted binding\nof alx predicted targets\n(neuron)",
  xlab = "DEGs in alx K.D. (Akheralie et al.)",
  ylab = "Weighted binding",
  las = 2,
  cex.main = .6,
  frame.plot = FALSE,
  outline = FALSE
)
stripchart(
  df_neu_alx$weighted_binding[f_neu]~df_neu_alx$deg[f_neu],
  col = deg_outl_col,
  method = "jitter",
  jitter=0.15,
  vertical = TRUE,
  pch = 20,
  cex=0.7,
  add = TRUE
)

## SECRETORY

set.seed(123)
non_deg_subsample_sec <- sample(df_sec_alx$tg[df_sec_alx$deg == "none"], 500)

f_sec <- which(df_sec_alx$tg %in% c(non_deg_subsample_sec,df_sec_alx$tg[df_sec_alx$deg != "none"]))

deg_cols <- setNames(c("#c4c4c4","#6428C8"),c("none","down"))
deg_outl_col <- deg_cols
deg_outl_col <- darken(deg_outl_col,.7)
deg_outl_col[1] <- alpha(deg_outl_col[1],.2)
deg_outl_col[c(2,3,4)] <- alpha(deg_outl_col[c(2,3,4)],.25)

set.seed(1234)
boxplot(
  df_sec_alx$prob~df_sec_alx$deg,
  col = deg_cols, border = darken(deg_cols,.7),
  main = "Interaction score\nof alx predicted targets\n(secretory)",
  xlab = "DEGs in alx K.D. (Akheralie et al.)",
  ylab = "Interaction score",
  las = 2,
  cex.main = .6,
  frame.plot = FALSE,
  outline = FALSE
)
stripchart(
  df_sec_alx$prob[f_sec]~df_sec_alx$deg[f_sec],
  col = deg_outl_col,
  method = "jitter",
  jitter=0.15,
  vertical = TRUE,
  pch = 20,
  cex=0.7,
  add = TRUE
)

set.seed(1234)
boxplot(
  df_sec_alx$weighted_binding~df_sec_alx$deg,
  col = deg_cols, border = darken(deg_cols,.7),
  main = "Weighted binding\nof alx predicted targets\n(secretory)",
  xlab = "DEGs in alx K.D. (Akheralie et al.)",
  ylab = "Weighted binding",
  las = 2,
  cex.main = .6,
  frame.plot = FALSE,
  outline = FALSE
)
stripchart(
  df_sec_alx$weighted_binding[f_sec]~df_sec_alx$deg[f_sec],
  col = deg_outl_col,
  method = "jitter",
  jitter=0.15,
  vertical = TRUE,
  pch = 20,
  cex=0.7,
  add = TRUE
)
par(mfrow = c(1,1))
dev.off()
```


## Saving all data

```{r}
save_to_excel(
  list_of_objects = 
    list(
      akheralieetal_bowtie2_counts = counts(dds,normalized = FALSE),
      akheralieetal_samples_table = sampletable,
      akheralieeatal_results = as.data.frame(res),
      alx_diffreg_norm_counts = assay(vsd)
    ),
  output_file_name =  paste0("./outputs/",fcha(),"_alx_akheralieetal_reanalysis.xlsx"),
  rowNames = TRUE
)
```

## Identifying potential candidates to validate secretory role

Below is the code (rough draft) that was used to manually inspect the feature plots of potential candidate genes.

The initial main criteria were:

 - Having a predicted weighted binding of 0.9 or higher with alx3-1 in the secretory network of ANANSE
 - Belonging to any of the secretory modules from our WGCNA analysis
 
Additional criteria:
 - Appears as differentially expressed in the knockdown RNA-Seq of alx3-1.
 
Eventually we saw that h1SMcG0000140 was among the best candidates, and in addition it has an in situ already from Fincher et al., 2018. So, we decided for h1SMcG0000140.


```{r, eval = FALSE}
candidate_insitus_targets_alx <-
  df_sec_alx$tg[
    df_sec_alx$weighted_binding > .9 &
    # df_sec_alx$prob > quantile(df_sec_alx$prob[df_sec_alx$deg=="down"],.95) &
    df_sec_alx$tg %in% smed_id_module$id[smed_id_module$module %in% smed_modules_table$newname[grep("secretory",smed_modules_table$celltypes)] ] &
    df_sec_alx$tg %in% alx_down
      ] # h1SMcG0000140, h1SMcG0017255, h1SMcG0021964

for (i in candidate_insitus_targets_alx){
  j = paste0("graphics/candidate_sec/candidate_sec_",i,".png")

  f <-FeaturePlot(
    scdata,
    features = i,
    cols = c("#DAE7F2","#531ccb"),
    order = TRUE,
    pt.size = 1
  )+ggtitle(i)

  png(j, wi = 512, he = 512)
  print(f)
  dev.off()

  message(i)
}

# c("h1SMcG0020662","h1SMcG0020160","h1SMcG0009505","h1SMcG0000426","h1SMcG0022558","h1SMcG0022546")

# mirar si 140 is high interactor en pou62 network of secretory
df_sec_pou <- sec_nw[sec_nw$tf == "h1SMcG0013730", -1]

# a vector of which is the cell type with highest expression for each gene (using our pseudobulk data).
smed_maxc <- setNames(apply(smed_counts_norm_cw, 1, function(x){names(x[x==max(x)])}),rownames(smed_counts_norm_cw))
maxc_sec <- names(smed_maxc[grep("secretory",smed_maxc)])
maxc_neu <- names(smed_maxc[smed_maxc %in% grep("neuron", colnames(smed_counts), va = T) ])

# DEGs alx
df_sec_alx$deg <-
  factor(ifelse(
    df_sec_alx$tg %in% alx_down #&
    # df_sec_alx$tg %in% maxc_sec
    , "down","none"
  ), levels = c("none","down"))

df_sec_alx$rank_prob <- relativise(rank(df_sec_alx$prob, ties.method = "min"))
df_sec_alx$rank_wb <- relativise(rank(df_sec_alx$weighted_binding, ties.method = "min"))


grep_candidates_insitus_targets_alx_2 <-
  c("0021964", "0028625", "0028112", "0028110", "0026716", "0021653", "0017439",
    "0015757", "0015756", "0005115", "0000205", "0000199", "0022571", "0022566",
    "0022558", "0022554", "0022547", "0022546", "0022447", "0021159", "0020934",
    "0020673", "0020600", "0019426", "0019147", "0018836", "0017915", "0017894",
    "0017787", "0017757", "0017756", "0017754", "0017429", "0017255", "0016534",
    "0016478", "0013672", "0009505", "0009304", "0009291", "0006020", "0005848",
    "0004969", "0002054", "0001554", "0000480", "0000359", "0000140", "0022089"
    )

c2 <- unname(unlist(sapply(grep_candidates_insitus_targets_alx_2,function(x){grep(x,candidate_insitus_targets_alx,value=TRUE)})))

c2_ <- c2[c2]

for (i in unname(unlist(sapply(grep_candidates_insitus_targets_alx_2,function(x){grep(x,candidate_insitus_targets_alx,value=TRUE)}))) ){

  s <- which(rosetta_ext$gene == i)

  name <- paste(i,rosetta_ext$Preferred_name[s],rosetta_ext$PF_name[s],rosetta_ext$PFAMs[s],rosetta_ext$name[s],sep="\n")

  j = paste0("graphics/candidate_sec/candidate_sec_2_",i,".png")

  f <-FeaturePlot(
    scdata,
    features = i,
    cols = c("#DAE7F2","#531ccb"),
    order = TRUE,
    pt.size = 1
    )+
    ggtitle(name)+
    theme(plot.title = element_text(size = 10, face = "bold"))

  png(j, wi = 512, he = 512)
  print(f)
  dev.off()

  message(i)
}
```

## Ploting in situ data

Here is the code used to plot the h1SMcG0000140 WISH quantification of number of cells per mm2 of regenerated tisue, per animal, per condition (control vs alx3-1 knockdown).

```{r}
nocells <- read.delim2("data/knockdown_data/alx_knockdown_WISH_nocells.tsv", dec = ".", header = TRUE)
nocells$condition = factor(nocells$condition, levels = c("gfp(RNAi)","alx3-1(RNAi)"))
alx_cols <- setNames(c("orange","purple"),c("gfp(RNAi)","alx3-1(RNAi)"))
wp <- wilcox.test(nocells$cells.per.mm2 ~ nocells$condition, alternative = "greater")$p.value
tp <- t.test(nocells$cells.per.mm2 ~ nocells$condition, alternative = "greater")$p.value
ymax <- round(max(nocells$cells.per.mm2)+ max(nocells$cells.per.mm2)*0.1)

set.seed(7002)
p_wish <-
  nocells %>%
  ggplot(
    mapping = 
      aes( 
        x = condition,
        y = cells.per.mm2,
        color = condition,
        fill = condition
      )
  )+
  geom_boxplot(outlier.shape = NA)+
  geom_jitter(size = 2, alpha = .75)+
  scale_fill_manual(values = lighten(alx_cols, .3) )+
  scale_color_manual(values = darken(alx_cols, .5) )+
  theme_classic()+
  theme(legend.position = "none")+
  ylim(0,ymax)+
  ylab("number of cells/mm^2 in\nregenerating area (1/6th of animal length)")+
  ggtitle("number of h1SMcG0000140 cells\nin alx3-1 knockdown")

pdf("graphics/alx_wish_boxplot.pdf",wi = 2.5, he = 5)
set.seed(7002)
print(p_wish)
dev.off()

print(p_wish)
```

## The End!

```{r}
sessionInfo()
```
